---
title: "Figure_brain"
author: "Kimberly Olney"
date: "06/15/2022"
output: html_document
---

Main figure showing brain results

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = ".")
```

May to unload all packages 
lapply(paste('package:',names(sessionInfo()$otherPkgs),sep=""),detach,character.only=TRUE,unload=TRUE)
```{r packages}
library(gprofiler2)
library(DESeq2)
library(enrichplot)
library(DOSE) 
library(plyr)
library(ggplot2)
require(gridExtra)
library(forcats)
library(UpSetR)
library(ComplexUpset)
library(ggplot2movies)
library(gtools)
library(ggnewscale)
library(ggrepel)
library(patchwork)
require(gridExtra)
library(ggpubr)
library(scales)
library("wiggleplotr")
library("dplyr")
library("GenomicRanges")
library("GenomicFeatures")
library("biomaRt")
library(tidyverse)
library(sleuth)
library(ggtranscript)
library(magrittr)
library(dplyr)
```

# User defined variables
```{r set_variables}
tissue <- "Brain"
control <- "Saline"
treatment <- "LPS"
control_color <- "gray65"
treatment_color <- "purple"
pathToRef <- c("/research/labs/neurology/fryer/projects/references/pig/ensembl_v7/")
```
# Save functions
These functions with help simultaneously save plots as a png, pdf, and tiff 
file.
```{r}
saveToPDF <- function(...) {
    d = dev.copy(pdf,...)
    dev.off(d)
}

saveToPNG <- function(...) {
    d = dev.copy(png,...)
    dev.off(d)
}
```

# Binary list shared gene function
```{r}
fromList <- function (input) {
  # Same as original fromList()...
  elements <- unique(unlist(input))
  data <- unlist(lapply(input, function(x) {
      x <- as.vector(match(elements, x))
      }))
  data[is.na(data)] <- as.integer(0)
  data[data != 0] <- as.integer(1)
  data <- data.frame(matrix(data, ncol = length(input), byrow = F))
  data <- data[which(rowSums(data) != 0), ]
  names(data) <- names(input)
  # ... Except now it conserves your original value names!
  row.names(data) <- elements
  return(data)
  }
```

# Gene level 
### Format dataframe
```{r}
tool = c("star")

treatment_vs_control <- read.table(
  paste0(
    "../../../results/",
    tool,
    "/DEGs/",
    treatment,
    "_",
    tissue,
    "_gene_DEGs_FDRq1.00.txt",
    sep = ""
  ),
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)

color_values <- vector()
max <- nrow(treatment_vs_control)

for (i in 1:max) {
  if (treatment_vs_control$adj.P.Val[i] < 0.05) {
    if (treatment_vs_control$logFC[i] > 0) {
      color_values <-
        c(color_values, 1) # 1 when logFC > 0 and FDRq < 0.05
    }
    else if (treatment_vs_control$logFC[i] < 0) {
      color_values <-
        c(color_values, 2) # 2 when logFC < 0 and FDRq < 0.05
    }
  }
  else{
    color_values <- c(color_values, 3) # 3 when FDRq >= 0.05
  }
}

treatment_vs_control$color_p0.05 <- factor(color_values)
up <- treatment_vs_control[treatment_vs_control$color_p0.05 == 1, ]
up10 <- up[1:5, ]
up10 <- subset(up10, gene_name != "TNFAIP3")

upFoldChange <-
  treatment_vs_control[treatment_vs_control$logFC > 4, ]
upFoldChange <- upFoldChange[4:12, ]
upFoldChange <- subset(upFoldChange, gene_name != "ENSSSCG00000006286")


down <- treatment_vs_control[treatment_vs_control$color_p0.05 == 2, ]
down <- subset(down, down$logFC < -1.5)
down10 <- down[1:3, ]

downFoldChange <-
  treatment_vs_control[treatment_vs_control$logFC < -3 &
                         treatment_vs_control$adj.P.Val < 0.05, ]
downFoldChange <- downFoldChange[5:10, ]

goi <-
  treatment_vs_control[treatment_vs_control$gene_name == "GATA2", ]
goi2 <-
  treatment_vs_control[treatment_vs_control$gene_name == "MARVELD2", ]

hadjpval <- (-log10(max(treatment_vs_control$P.Value[treatment_vs_control$adj.P.Val < 0.05],
                        na.rm = TRUE)))
```
### Volcano plot
```{r}
gene_volcano <-
  ggplot(data = treatment_vs_control,
         aes(
           x = logFC,
           y = -log10(P.Value),
           color = color_p0.05
         )) +  
  geom_point(alpha = 0.5, size = 1) +
  theme_bw() +  
  theme(legend.position = "none") + 
  scale_color_manual(values = c("red", "blue", "grey")) + 
  labs(x = expression(log[2](FC)),
       y = expression(-log[10] ~ "(" ~ italic("p") ~ "-value)")) +
       theme(plot.margin = margin(0.1, 0.2, 0, 0.2, "cm"), 
             axis.title.x = element_text(size = 6),
             axis.text.x = element_text(size = 6)) +
         theme(axis.title.y = element_text(size = 6),
               axis.text.y = element_text(size = 6)) +
         geom_hline(yintercept = hadjpval,
                    #  horizontal line
                    colour = "#000000",
                    linetype = "dashed") +
        ggtitle(paste0(tissue,"\nFDRq < 0.05")) +
         theme(plot.title = element_text(size = 7, margin = margin(0,0,0,0))) +
         geom_text_repel(
           data = up10,
           aes(
             x = logFC,
             y = -log10(P.Value),
             label = gene_name
           ),
           color = "maroon",
           fontface = "italic",
           size = 2,
           min.segment.length = unit(0, 'lines'),
          # nudge_x = 0,
         #  nudge_y = 0.5,
           max.overlaps = getOption("ggrepel.max.overlaps", default = 20)
         ) +
         geom_text_repel(
           data = upFoldChange,
           aes(
             x = logFC,
             y = -log10(P.Value),
             label = gene_name
           ),
           color = "maroon",
           fontface = "italic",
           size = 2,
           min.segment.length = unit(0, 'lines'),
           nudge_x = .6,
           max.overlaps = getOption("ggrepel.max.overlaps", default = 10)
         ) +
         geom_text_repel(
           data = down10,
           aes(
             x = logFC,
             y = -log10(P.Value),
             label = gene_name
           ),
           color = "navyblue",
           fontface = "italic",
           size = 2,
           min.segment.length = unit(0, 'lines'),
           nudge_x = -1,
           nudge_y = .25,
           max.overlaps = getOption("ggrepel.max.overlaps", default = 10)
         )  +
      #   geom_text_repel(
      #     data = downFoldChange,
      #     aes(
      #       x = logFC,
      #       y = -log10(P.Value),
      #       label = gene_name
      #     ),
       #    color = "navyblue",
       #    fontface = "italic",
        #   size = 2,
        #   min.segment.length = unit(0, 'lines'),
        #   nudge_x = -.5,
        #   nudge_y = -.5,
       #    max.overlaps = getOption("ggrepel.max.overlaps", default = 5)
       #  ) +
         geom_text_repel(
           data = goi,
           aes(
             x = logFC,
             y = -log10(P.Value),
             label = gene_name
           ),
           color = "navyblue",
           fontface = "italic",
           size = 2,
           nudge_x = -2,
           nudge_y = 0.5,
           min.segment.length = unit(0, 'lines'),
           max.overlaps = getOption("ggrepel.max.overlaps", default = 5)
         ) +
         geom_text_repel(
           data = goi2,
           aes(
             x = logFC,
             y = -log10(P.Value),
             label = gene_name
           ),
           color = "navyblue",
           fontface = "italic",
           size = 2,
           nudge_x = 2.65,
           nudge_y = .75,
           min.segment.length = unit(0, 'lines'),
           max.overlaps = getOption("ggrepel.max.overlaps", default = 5)
         ) +
         scale_y_continuous(breaks = seq(0, 12, by = 1), limits = c(0, 12)) +
         scale_x_continuous(breaks = seq(-8, 8, by = 2), limits = c(-8, 8))
gene_volcano
```
remove files no longer needed
```{r}
remove(
  down,
  down10,
  downFoldChange,
  goi,
  goi2,
  up,
  up10,
  upFoldChange,
  hadjpval,
  i,
  max,
  color_values
)
```

# Metascape Enrichment 
### Format the dataframes
```{r}
# read in enrichment analysis results
up_enrich_results <-
  read.delim(
    paste0(
      "../../../results/",
      tool,
      "/metascape/",
      tolower(tissue),
      "_up.txt"
    ),
    sep = "\t",
    header = TRUE
  )

down_enrich_results <-
  read.delim(
    paste0(
      "../../../results/",
      tool,
      "/metascape/",
      tolower(tissue),
      "_down.txt"
    ),
    sep = "\t",
    header = TRUE
  )
```

```{r}
# select the GO term IDs we want to show in the plot
GO_ID_up <-
  c(
    "1_Summary",
    "2_Summary",
    "3_Summary",
    "4_Summary",
    "5_Summary",
    "6_Summary",
    "7_Summary",
    "8_Summary",
    "9_Summary", 
    "10_Summary",
    "11_Summary",
    "12_Summary",
    "13_Summary",
    "14_Summary",
    "15_Summary",
    "16_Summary",
    "17_Summary",
    "18_Summary",
    "19_Summary", 
    "20_Summary"
  )
GO_ID_down <- c(
    "1_Summary",
    "2_Summary",
    "3_Summary",
    "4_Summary",
    "5_Summary",
    "6_Summary",
    "7_Summary",
    "8_Summary",
    "9_Summary", 
    "10_Summary",
    "11_Summary",
    "12_Summary",
    "13_Summary",
    "14_Summary",
    "15_Summary",
    "16_Summary",
    "17_Summary",
    "18_Summary",
    "19_Summary", 
    "20_Summary"
)

up_enrich_results_subset <-
  up_enrich_results[up_enrich_results$GroupID %in% GO_ID_up, ]
up_enrich_results_subset$Cluster <- c("up-regulated")
up_enrich_results_subset$Description <-
  factor(up_enrich_results_subset$Description,
         levels = up_enrich_results_subset$Description)
up_enrich_results_subset$Description <-
  fct_rev(up_enrich_results_subset$Description)

down_enrich_results_subset <-
  down_enrich_results[down_enrich_results$GroupID %in% GO_ID_down, ]
down_enrich_results_subset$Cluster <- c("down-regulated")
down_enrich_results_subset$Description <-
  factor(down_enrich_results_subset$Description,
         levels = down_enrich_results_subset$Description)
down_enrich_results_subset$Description <-
  fct_rev(down_enrich_results_subset$Description)

# get the number of genes in each summary
up_gene_count <-
  strsplit(as.character(up_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
up_gene_count_df <-
  data.frame(matrix(
    unlist(up_gene_count),
    nrow = length(up_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
up_enrich_results_subset$InTerm <- as.numeric(up_gene_count_df$X1)
up_enrich_results_subset$InList <- as.numeric(up_gene_count_df$X2)

down_gene_count <-
  strsplit(as.character(down_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
down_gene_count_df <-
  data.frame(matrix(
    unlist(down_gene_count),
    nrow = length(down_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
down_enrich_results_subset$InTerm <- as.numeric(down_gene_count_df$X1)
down_enrich_results_subset$InList <- as.numeric(down_gene_count_df$X2)

# combine together
up_and_down_enrich_results_subset <- rbind(up_enrich_results_subset, down_enrich_results_subset)

up_and_down_enrich_results_subset$Description <-
      gsub(
        "regulation of DNA-binding transcription factor activity",
        "reg of DNA-binding TF activity",
        up_and_down_enrich_results_subset$Description
      )
up_and_down_enrich_results_subset$Description <-
  factor(up_and_down_enrich_results_subset$Description, levels = unique(up_and_down_enrich_results_subset$Description)) 

up_and_down_enrich_results_subset$Description <- fct_rev(up_and_down_enrich_results_subset$Description)
```

remove files 
```{r}
remove(
  down_enrich_results,
  down_gene_count,
  down_gene_count_df,
  up_enrich_results,
  up_gene_count,
  up_gene_count_df
)
```

### Enrichment plot showing log10 P-value
#### down and up seperately
```{r}
down_enrich_plot <-
  ggplot(data = down_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster) +
  geom_bar(stat = "identity", aes(fill = LogP), width = .65) +
  theme(axis.title.x = element_text(size = 6),
        axis.text.x = element_text(size = 6)) +
  theme(axis.title.y = element_text(size = 6),
        axis.text.y = element_text(size = 6)) +
  theme_bw() +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = "Gene count", y = "") +
  #  xlim(0,60) +
  guides(fill = guide_legend(
    size = 1,
    title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)")
  )) +
  scale_fill_gradientn(
    colours = c("darkmagenta", "mediumorchid3", "grey"),
    guide = "legend",
    limits = c(-45,-1),
  ) +
  theme(legend.text = element_text(size = 6),
        legend.title = element_text(size = 6))
down_enrich_plot

up_enrich_plot <-
  ggplot(data = up_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster) +
  geom_bar(stat = "identity", aes(fill = LogP), width = .65) +
  theme(axis.title.x = element_text(size = 6),
        axis.text.x = element_text(size = 6)) +
  theme(axis.title.y = element_text(size = 6),
        axis.text.y = element_text(size = 6)) +
  theme(legend.text = element_text(size = 6)) +
  theme_bw() +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = "Gene count", y = "") +
  guides(fill = guide_legend(title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)"))) +
  scale_fill_gradientn(
    colours = c("darkmagenta", "mediumorchid3", "grey"),
    guide = "legend",
    limits = c(-45,-1)
  ) +
  theme(legend.text = element_text(size = 6),
        legend.title = element_text(size = 6))
up_enrich_plot
```

#### down and up together
```{r}
up_and_down_enrich_plot <-
  ggplot(data = up_and_down_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster, scales = "free") +
  geom_bar(stat = "identity", aes(fill = LogP), width = .7, position = position_dodge(width = .2)) +
  theme_bw() +
  ggtitle(paste0(tissue," enrichment summaries")) +
  labs(x = "Gene count", y = NULL) +
  guides(fill = guide_legend(title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)"))) +
  scale_fill_gradientn(
    colours = c("darkmagenta", "mediumorchid3", "grey"),
    guide = "legend",
    limits = c(-45,-1)
  ) +
    theme(strip.text = element_text(size = 6), 
          axis.text.y = element_text(size = 7),
          axis.text.x = element_text(size = 6), 
          axis.title.x = element_text(size = 6), 
          legend.text = element_text(size = 6),
          legend.title = element_text(size = 6),
          legend.margin=margin(0,0.5,0,0),
          legend.box.margin=margin(-10,-2,-10,-7.5), 
          plot.margin = margin(0.1, 0.2, 0, 0.2, "cm"), 
          plot.title = element_text(size = 7, hjust = -3.25, vjust=0, margin = margin(0,0,0,0)))
up_and_down_enrich_plot

addSmallLegend <- function(myPlot, pointSize = 3, textSize = 6, spaceLegend = .5) {
    myPlot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}

# Apply on original plot
up_and_down_enrich_plot <- addSmallLegend(up_and_down_enrich_plot)
```

```{r}
remove(up_enrich_plot,
       down_enrich_plot)
```
## Combine volcano and enrichment plot 
```{r}
row1 <- ggarrange(
  gene_volcano,
  up_and_down_enrich_plot,
  ncol = 2,
  labels = c("a)", "b)"),
  widths = c(2.25, 4.75), 
  font.label = list(size = 8)
  )
row1

#path <- paste0("../../../results/manuscript_figures/Figure_",
#         tolower(tissue),
 #        "row1_test")
#saveToPDF(paste0(path, ".pdf"), width = 7, height = 3)
```
# CPM data
```{r}
# read in counts per million
cpm <- read.delim(paste0("../../../results/", tool, "/CPM/", treatment, "_",tolower(tissue),
               "_gene_CPM_expression.txt"))
# then log2 transform 
lcpm <- edgeR::cpm(cpm, log = TRUE)
# save ouput
#write.table(lcpm, paste0("../../../results/", tool, "/CPM/", treatment, "_",tolower(tissue),
  #             "_gene_log2CPM_expression.txt"), quote = FALSE, sep = "\t")
```
### Fold change information 
```{r}
up_regulated <- subset(
  treatment_vs_control,
  treatment_vs_control$adj.P.Val <= 0.05 &
    treatment_vs_control$logFC > 0
)
down_regulated <- subset(
  treatment_vs_control,
  treatment_vs_control$adj.P.Val <= 0.05 &
    treatment_vs_control$logFC <= 0
)

# sort by log2FC and adjusted p-val
up_regulated_sort <-
  up_regulated[order(-up_regulated$logFC, up_regulated$adj.P.Val), ]
down_regulated_sort <-
  down_regulated[order(down_regulated$logFC, down_regulated$adj.P.Val), ]

# get the lcpm information for those genes of interest
up_regulated_lcpm <- subset(lcpm, row.names(lcpm) %in%
                              up_regulated_sort$gene_id)
down_regulated_lcpm <- subset(lcpm, row.names(lcpm) %in%
                                down_regulated_sort$gene_id)

# reform the data
up_regulated_sort_df <- up_regulated_lcpm %>%
  as.data.frame() %>%
  rownames_to_column("gene_id") %>%
  pivot_longer(-c(gene_id), names_to = "samples", values_to = "counts") %>%
  mutate(samples = fct_relevel(samples, colnames(up_regulated_lcpm)))

down_regulated_sort_df <- down_regulated_lcpm %>%
  as.data.frame() %>%
  rownames_to_column("gene_id") %>%
  pivot_longer(-c(gene_id), names_to = "samples", values_to = "counts") %>%
  mutate(samples = fct_relevel(samples, colnames(down_regulated_lcpm)))

# lock in gene order which is sorted by adjusted p-value
up_regulated_sort_df$gene_id <-
  factor(up_regulated_sort_df$gene_id,
         levels = unique(up_regulated_sort_df$gene_id))
up_regulated_sort_df$gene_id <-
  fct_rev(up_regulated_sort_df$gene_id)
down_regulated_sort_df$gene_id <-
  factor(down_regulated_sort_df$gene_id,
         levels = unique(down_regulated_sort_df$gene_id))
down_regulated_sort_df$gene_id <-
  fct_rev(down_regulated_sort_df$gene_id)

DEG_df <-
  rbind(up_regulated_sort_df, down_regulated_sort_df)

DEG_df$Condition <-
  ifelse(grepl("LPS", DEG_df$samples),
         "LPS",
         "Control")

DEG_df$samples <- factor(
  DEG_df$samples,
  levels = c(
    "X3.Control",
    "X4.Control",
    "X6.Control",
    "X7.Control",
    "X10.Control",
    "X11.Control",
    "X2.LPS",
    "X8.LPS",
    "X12.LPS",
    "X14.LPS"
  )
)
DEG_sort <- rbind(up_regulated_sort, down_regulated_sort)
names(DEG_sort)[names(DEG_sort) == "gene_name"] <- "gene"

```
remove files
```{r}
remove(
  cpm,
  lcpm,
  down_regulated,
  down_regulated_lcpm,
  down_regulated_sort,
  down_regulated_sort_df,
  up_regulated,
  up_regulated_lcpm,
  up_regulated_sort,
  up_regulated_sort_df
)
```

### Genes in enrichment plot all
## up 
 [1] Cytokine Signaling in Immune system                    
 [2] cellular response to cytokine stimulus                 
 [3] TNF signaling pathway                                  
 [4] response to virus                                      
 [5] positive regulation of cytokine production             
 [6] regulation of defense response                         
 [7] positive regulation of cell death                      
 [8] hematopoietic or lymphoid organ development            
 [9] regulation of response to biotic stimulus              
[10] Orexin receptor pathway                                
[11] regulation of cell adhesion                            
[12] positive regulation of cell migration                  
[13] Network map of SARS-CoV-2 signaling pathway            
[14] vasculature development                                
[15] regulation of angiogenesis                             
[16] response to tumor necrosis factor                      
[17] PID P53 DOWNSTREAM PATHWAY                             
[18] PID AP1 PATHWAY                                        
[19] VEGFA-VEGFR2 signaling pathway                         
[20] regulation of DNA-binding transcription factor activity
```{r}
up_enrich_df <-
  up_enrich_results_subset[, c("GroupID", "Term", "Description", "Symbols")]
up_enrich_df_melt <- reshape2::melt(up_enrich_df)
up_gene_char <- str_split(up_enrich_df_melt$Symbols, ",")

up_gene_pathway <-
  data.frame(gene = unlist(up_gene_char),
             value = rep(up_enrich_df$Description, lengths(up_gene_char)))
up_pathways_levels <- levels(fct_rev(up_gene_pathway$value))

up_enrich_df <-
  up_enrich_results_subset[, c("GroupID", "Term", "Description", "Symbols")]
up_enrich_df_melt <- reshape2::melt(up_enrich_df)
up_gene_char <- str_split(up_enrich_df_melt$Symbols, ",")

up_gene_pathway <-
  data.frame(gene = unlist(up_gene_char),
             value = rep(up_enrich_df$Description, lengths(up_gene_char)))
up_pathways_levels <- levels(fct_rev(up_gene_pathway$value))

up_input <- list(
  "Cytokine Signaling in Immune system" = up_gene_char[[1]],
  "positive regulation of cell death" = up_gene_char[[7]],
  "positive regulation of cell migration" = up_gene_char[[12]],
  "vasculature development" = up_gene_char[[14]],
  "regulation of DNA-binding transcription factor activity" = up_gene_char[[20]] 
)
up_data <- fromList(up_input)
up_data$gene <- row.names(up_data)
up_data_melt <- reshape2::melt(up_data)

# get the fold change value for those genes
up_df <- merge(up_data_melt, DEG_sort,
                       by = "gene")
# order by fold change 
up_df <- up_df[order(-up_df$logFC, -up_df$adj.P.Val),]

# subset by GO termes
GO_1 <-
  subset(up_df,
         up_df$variable == "Cytokine Signaling in Immune system"
         & up_df$value == 1)
GO_1_top <- head(GO_1, 15)

GO_2 <-
  subset(up_df,
         up_df$variable == "positive regulation of cell death"
         & up_df$value == 1)
GO_2_top <- head(GO_2, 15)

GO_3 <-
  subset(up_df,
         up_df$variable == "positive regulation of cell migration"
         & up_df$value == 1)
GO_3_top <- head(GO_3, 15)

GO_4 <-
  subset(up_df,
         up_df$variable == "vasculature development"
         & up_df$value == 1)
GO_4_top <- head(GO_4, 15)

GO_5 <-
  subset(up_df,
         up_df$variable == "regulation of DNA-binding transcription factor activity"
         & up_df$value == 1)
GO_5_top <- head(GO_5, 15)
```

#### plot lcpm GO heatmaps
##### up 1
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_1_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_1_top$gene_id)
GO_1_DEG <- merge(GO_1_DEG, GO_1_top, by = "gene_id", all.x = TRUE)
GO_1_DEG$gene <- factor(GO_1_DEG$gene,
                          levels = unique(GO_1_top$gene))
GO_1_DEG$gene <- fct_rev(GO_1_DEG$gene)
# heatmap 
up_GO_1_lcpm_heat <- ggplot(data = GO_1_DEG, aes(x = samples, y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ Condition, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 4, 8, 12, 16),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("Cytokine Signaling in 
Immune system") 
up_GO_1_lcpm_heat
```
##### up 2
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_2_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_2_top$gene_id)
GO_2_DEG <- merge(GO_2_DEG, GO_2_top, by = "gene_id", all.x = TRUE)
GO_2_DEG$gene <- factor(GO_2_DEG$gene,
                          levels = unique(GO_2_top$gene))
GO_2_DEG$gene <- fct_rev(GO_2_DEG$gene)

# heatmap 
up_GO_2_lcpm_heat <- ggplot(data = GO_2_DEG, aes(x = samples, y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ Condition, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 4, 8, 12, 16),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("positive regulation 
of cell death")
up_GO_2_lcpm_heat
```
##### up 3
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_3_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_3_top$gene_id)
GO_3_DEG <- merge(GO_3_DEG, GO_3_top, by = "gene_id", all.x = TRUE)
GO_3_DEG$gene <- factor(GO_3_DEG$gene,
                          levels = unique(GO_3_top$gene))
GO_3_DEG$gene <- fct_rev(GO_3_DEG$gene)

# heatmap 
up_GO_3_lcpm_heat <- ggplot(data = GO_3_DEG, aes(x = samples, y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ Condition, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 4, 8, 12, 16),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("positive regulation 
of cell migration")
up_GO_3_lcpm_heat
```
##### up 4
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_4_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_4_top$gene_id)
GO_4_DEG <- merge(GO_4_DEG, GO_4_top, by = "gene_id", all.x = TRUE)

GO_4_DEG$gene <- factor(GO_4_DEG$gene,
                          levels = unique(GO_4_top$gene))
GO_4_DEG$gene <- fct_rev(GO_4_DEG$gene)

# heatmap 
up_GO_4_lcpm_heat <- ggplot(data = GO_4_DEG, aes(x = samples, y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ Condition, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 4, 8, 12, 16),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("vasculature 
development")
up_GO_4_lcpm_heat
```

##### up 5
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_5_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_5_top$gene_id)
GO_5_DEG <- merge(GO_5_DEG, GO_5_top, by = "gene_id", all.x = TRUE)

GO_5_DEG$gene <- factor(GO_5_DEG$gene,
                          levels = unique(GO_5_top$gene))
GO_5_DEG$gene <- fct_rev(GO_5_DEG$gene)

# heatmap 
up_GO_5_lcpm_heat <- ggplot(data = GO_5_DEG, aes(x = samples, y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ Condition, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 4, 8, 12, 16),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("reg of DNA-binding 
TF activity")
up_GO_5_lcpm_heat
```

clean up
```{r, eval = FALSE}
remove(
  up_data,
  up_data_melt,
  up_df,
  up_enrich_df,
  up_enrich_df_melt,
  up_enrich_results_subset,
  up_gene_char,
  up_gene_pathway,
  up_input,
  GO_1,
  GO_1_DEG,
  GO_1_top,
  GO_2,
  GO_2_DEG,
  GO_2_top,
  GO_3,
  GO_3_DEG,
  GO_3_top,
  GO_4,
  GO_4_DEG,
  GO_4_top,
  GO_5, 
  GO_5_DEG,
  GO_5_top
)
```
## down
 [1] immune system development                      
 [2] organic acid transport                         
 [3] vitamin transport                              
 [4] regulation of actin cytoskeleton organization  
 [5] embryonic organ morphogenesis                  
 [6] stem cell proliferation                        
 [7] lipid transport                                
 [8] Malignant pleural mesothelioma                 
 [9] regulation of actin cytoskeleton reorganization
[10] tight junction assembly                        
[11] semicircular canal development                 
[12] Hematopoietic stem cell differentiation        
[13] blood vessel endothelial cell migration        
[14] gland development                              
[15] Nuclear receptors meta-pathway                 
[16] positive regulation of angiogenesis            
[17] regulation of animal organ morphogenesis       
[18] sulfur compound catabolic process              
[19] morphogenesis of an epithelium                 
[20] regulation of binding 
```{r}
down_enrich_results_subset$Description
down_enrich_df <-
  down_enrich_results_subset[, c("GroupID", "Term", "Description", "Symbols")]
down_enrich_df_melt <- reshape2::melt(down_enrich_df)
down_gene_char <- str_split(down_enrich_df_melt$Symbols, ",")

down_gene_pathway <-
  data.frame(gene = unlist(down_gene_char),
             value = rep(down_enrich_df$Description, lengths(down_gene_char)))
down_pathways_levels <- levels(fct_rev(down_gene_pathway$value))

down_input <- list(
  "immune system development" = down_gene_char[[1]],
  "organic acid transport" = down_gene_char[[2]],
  "tight junction assembly" = down_gene_char[[10]], 
  "blood vessel endothelial cell migration" = down_gene_char[[13]],
  "positive regulation of angiogenesis" = down_gene_char[[16]]
)
down_data <- fromList(down_input)
down_data$gene <- row.names(down_data)
down_data_melt <- reshape2::melt(down_data)


# get the fold change value for those genes
down_df <- merge(down_data_melt, DEG_sort,
                       by = "gene")
# order by fold change 
down_df <- down_df[order(down_df$logFC, -down_df$adj.P.Val),]

# subset by GO termes
GO_1 <-
  subset(down_df,
         down_df$variable == "immune system development"
         & down_df$value == 1)
GO_1_top <- head(GO_1, 15)

GO_2 <-
  subset(down_df,
         down_df$variable == "organic acid transport"
         & down_df$value == 1)
GO_2_top <- head(GO_2, 15)

GO_3 <-
  subset(down_df,
         down_df$variable == "tight junction assembly"
         & down_df$value == 1)
GO_3_top <- head(GO_3, 15)

GO_4 <-
  subset(down_df,
         down_df$variable == "blood vessel endothelial cell migration"
         & down_df$value == 1)
GO_4_top <- head(GO_4, 15)

GO_5 <-
  subset(down_df,
         down_df$variable == "positive regulation of angiogenesis"
         & down_df$value == 1)
GO_5_top <- head(GO_5, 15)
```
#### plot lcpm GO heatmaps
##### down 1
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_1_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_1_top$gene_id)
GO_1_DEG <- merge(GO_1_DEG, GO_1_top, by = "gene_id", all.x = TRUE)
GO_1_DEG$gene <- factor(GO_1_DEG$gene,
                          levels = unique(GO_1_top$gene))
GO_1_DEG$gene <- fct_rev(GO_1_DEG$gene)

# heatmap 
down_GO_1_lcpm_heat <- ggplot(data = GO_1_DEG, aes(x = samples, y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ Condition, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 4, 8, 12, 16),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
  ggtitle("immune system 
development") 
down_GO_1_lcpm_heat
```
##### down 2
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_2_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_2_top$gene_id)
GO_2_DEG <- merge(GO_2_DEG, GO_2_top, by = "gene_id", all.x = TRUE)
GO_2_DEG$gene <- factor(GO_2_DEG$gene,
                          levels = unique(GO_2_top$gene))
GO_2_DEG$gene <- fct_rev(GO_2_DEG$gene)

# heatmap 
down_GO_2_lcpm_heat <- ggplot(data = GO_2_DEG, aes(x = samples, y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ Condition, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 4, 8, 12, 16),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("organic acid 
transport")
down_GO_2_lcpm_heat
```
##### down 3
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_3_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_3_top$gene_id)
GO_3_DEG <- merge(GO_3_DEG, GO_3_top, by = "gene_id", all.x = TRUE)
GO_3_DEG$gene <- factor(GO_3_DEG$gene,
                          levels = unique(GO_3_top$gene))
GO_3_DEG$gene <- fct_rev(GO_3_DEG$gene)

# heatmap 
down_GO_3_lcpm_heat <- ggplot(data = GO_3_DEG, aes(x = samples, y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ Condition, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 4, 8, 12, 16),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("tight junction
assembly")
down_GO_3_lcpm_heat
```
##### down 4
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_4_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_4_top$gene_id)
GO_4_DEG <- merge(GO_4_DEG, GO_4_top, by = "gene_id", all.x = TRUE)
GO_4_DEG$gene <- factor(GO_4_DEG$gene,
                          levels = unique(GO_4_top$gene))
GO_4_DEG$gene <- fct_rev(GO_4_DEG$gene)

# heatmap 
down_GO_4_lcpm_heat <- ggplot(data = GO_4_DEG, aes(x = samples, y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ Condition, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 4, 8, 12, 16),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("blood vessel 
endothelial 
cell migration")
down_GO_4_lcpm_heat
```
##### down 5
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_5_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_5_top$gene_id)
GO_5_DEG <- merge(GO_5_DEG, GO_5_top, by = "gene_id", all.x = TRUE)
GO_5_DEG$gene <- factor(GO_5_DEG$gene,
                          levels = unique(GO_5_top$gene))
GO_5_DEG$gene <- fct_rev(GO_5_DEG$gene)

# heatmap 
down_GO_5_lcpm_heat <- ggplot(data = GO_5_DEG, aes(x = samples, y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ Condition, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 4, 8, 12, 16),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0, 0.2, 0, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("positive regulation 
of angiogenesis")
down_GO_5_lcpm_heat
```

### legend only
```{r}
# Extract the legend. Returns a gtable
# lock in gene order which is sorted by adjusted p-value
GO_1_DEG <- subset(DEG_df, DEG_df$gene_id %in% GO_1_top$gene_id)
GO_1_DEG <- merge(GO_1_DEG, GO_1_top, by = "gene_id", all.x = TRUE)
GO_1_DEG$gene <- factor(GO_1_DEG$gene,
                          levels = unique(GO_1_top$gene))
GO_1_DEG$gene <- fct_rev(GO_1_DEG$gene)

# heatmap 
legend_data <- ggplot(data = GO_1_DEG, aes(x = samples, y = gene)) +
  geom_tile(aes(fill = counts)) +
  ggplot2::facet_grid(~ Condition, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "#FFFFCCFF",
    mid = "#FD8D3CFF",
    high = "#800026FF",
    midpoint = 6,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 4, 8, 12, 16),
    limits = c(0,16),
    labels=c(0, 4, 8, 12, 16), 
    name = expression(log[2](CPM))
  ) +
  theme(
    legend.position = "top", 
    legend.text = element_text(size =6), 
    legend.title = element_text (size = 6), 
    legend.margin=margin(0,0,1,0),
  #  legend.direction = "horizontal", 
    legend.key.size = unit(0.5, "cm"))

legend_data <- addSmallLegend(legend_data)

leg <- get_legend(legend_data)
# Convert to a ggplot and print
legend_heat_bar <- as_ggplot(leg)
legend_heat_bar
```
clean up
```{r, eval =FALSE}
remove(
  down_data,
  down_data_melt,
  down_df,
  down_enrich_df,
  down_enrich_df_melt,
  down_enrich_results_subset,
  down_gene_char,
  down_gene_pathway,
  down_input,
  GO_1,
  GO_1_DEG,
  GO_1_top,
  GO_2,
  GO_2_DEG,
  GO_2_top,
  GO_3,
  GO_3_DEG,
  GO_3_top,
  GO_4,
  GO_4_DEG,
  GO_4_top,
  GO_5, 
  GO_5_DEG,
  GO_5_top
)
```

# Gene-level combine all plots 
```{r}
row2 <-
  ggarrange(
    NULL,
    legend_heat_bar,
    NULL,
    NULL,
    ncol = 4,
    labels = c("up-regulated", "", "", ""),
    font.label = list(size = 7, color = "red")
  )

row3 <- ggarrange(
  up_GO_1_lcpm_heat,
  up_GO_2_lcpm_heat,
  up_GO_3_lcpm_heat,
  up_GO_4_lcpm_heat,
  up_GO_5_lcpm_heat,
  ncol = 5,
  labels = c("c)", "d)", "e)", "f)", "g)"),
  font.label = list(size = 8)
  )

row4 <-
  ggarrange(
    NULL,
    NULL,
    NULL,
    NULL,
    ncol = 4,
    labels = c("down-regulated", "", "", ""),
    font.label = list(size = 7, color = "blue")
  )

row5 <- ggarrange(
  down_GO_1_lcpm_heat,
  down_GO_2_lcpm_heat,
  down_GO_3_lcpm_heat,
  down_GO_4_lcpm_heat,
  down_GO_5_lcpm_heat,
  ncol = 5,
  labels = c("h)", "i)", "j)", "k)", "l)"),
  font.label = list(size = 8)
  )

combind <-
  ggarrange(
    row1,
    row2,
    row3,
    row4,
    row5,
    nrow = 5,
    heights = c(1.65, 0.1, 1, 0.1, 1)
  )
combind

path <- paste0("../../../results/manuscript_figures/Figure_",
               tolower(tissue),
               "_gene_level_only")
saveToPDF(paste0(path, ".pdf"), width = 7.08, height = 8.66)
```
