---
title: "Figure_mouse_and_pig_brain"
author: "Kimberly Olney"
date: "09/12/2022"
output: html_document
---

Main figure showing brain results

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = ".")
```

```{r, message = FALSE, warning = FALSE, echo = FALSE, include = FALSE}
library(geneName)
library(plotly)
library(tidyverse)
library(htmlwidgets)
library(ggvenn)
library(dplyr)
library(gprofiler2)
library(stats)
library(UpSetR)
library(ComplexUpset)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE, include = FALSE}
saveToPDF <- function(...) {
  d = dev.copy(pdf, ...)
  dev.off(d)
}

addSmallLegend <- function(myPlot, pointSize = 3, textSize = 6, spaceLegend = .5) {
    myPlot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}
```
# Binary list shared gene function
```{r}
fromList <- function (input) {
  # Same as original fromList()...
  elements <- unique(unlist(input))
  data <- unlist(lapply(input, function(x) {
      x <- as.vector(match(elements, x))
      }))
  data[is.na(data)] <- as.integer(0)
  data[data != 0] <- as.integer(1)
  data <- data.frame(matrix(data, ncol = length(input), byrow = F))
  data <- data[which(rowSums(data) != 0), ]
  names(data) <- names(input)
  # ... Except now it conserves your original value names!
  row.names(data) <- elements
  return(data)
  }
```
# Read in DEG tables
Note that gene names have already been converted to the human ortholog\
see correlation_analysis_mouse_pig_brain.Rmd for more details. 
```{r}
pig_brain_FDRq1.00_orth_1n <- read.delim("../../../mouse_comparison/results/pig_brain_FDRq1.00_orth_1n.txt",
  sep = "\t",
  header = T
)

mouse_limma_FDRq1.00_orth_1n <- read.delim(
  "../../../mouse_comparison/results/mouse_limma_FDRq1.00_orth_1n.txt",
  sep = "\t",
  header = T
)
```

# AdjPval 1.00 Find shared and unique DEGs between mouse and pig
```{r}
# rename GeneName to gene_name to merge data sets and find intersections
mouse_and_pg_df <-
  merge(mouse_limma_FDRq1.00_orth_1n, pig_brain_FDRq1.00_orth_1n, by = "ortholog_name")

shared_gene_name <-
  intersect(mouse_limma_FDRq1.00_orth_1n$ortholog_name,
            pig_brain_FDRq1.00_orth_1n$ortholog_name)
uniqueToMouse_gene_name <-
  setdiff(mouse_limma_FDRq1.00_orth_1n$ortholog_name,
          pig_brain_FDRq1.00_orth_1n$ortholog_name)
uniqueToPig_gene_name <-
  setdiff(pig_brain_FDRq1.00_orth_1n$ortholog_name,
          mouse_limma_FDRq1.00_orth_1n$ortholog_name)

shared_adjPval_1 <-
  mouse_and_pg_df[mouse_and_pg_df$ortholog_name %in% shared_gene_name,]
uniqueToMouse <-
  mouse_and_pg_df[mouse_and_pg_df$ortholog_name %in% uniqueToMouse_gene_name,]
uniqueToPig <-
  mouse_and_pg_df[mouse_and_pg_df$ortholog_name %in% uniqueToPig_gene_name,]


# Venn diagram to see total DEGs between mouse and pig
x = list(mouse = mouse_limma_FDRq1.00_orth_1n$ortholog_name,
         pig = pig_brain_FDRq1.00_orth_1n$ortholog_name)

mouse_pig_venn <- ggvenn(
  x,
  fill_color = c("#EFC000FF", "purple"),
  stroke_size = 2,
  set_name_size = 6
)
mouse_pig_venn
```

## correlation plot - all
```{r}
corr_plot_adjPval_1 <- ggplot(data = shared_adjPval_1, aes(x = logFC.x, y = logFC.y, text = paste(gene_name))) +
   annotate(
    "rect",
    xmin = 0,
    xmax = 9,
    ymin = 9,
    ymax = 0,
    fill = "lightpink3",
    alpha = .5
  ) +  annotate(
    "rect",
    xmin = 0,
    xmax = -4,
    ymin = 0,
    ymax = -4,
    fill = "cadetblue3",
    alpha = .5
  ) +
  geom_abline(color = "gray40") +
  geom_point(size = 1) + 
  theme_bw() +
  theme(plot.title = element_text(size = 7, margin = margin(0,0,0,0))) +
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size = 6),
        axis.text.x = element_text(size = 6)) +
  theme(axis.title.y = element_text(size = 6),
        axis.text.y = element_text(size = 6)) +
  labs(
  title = "All common genes between mouse and pig brains
adjusted p.value < 1.00",
  x = expression(paste("mouse ", log[2](FC))),
  y = expression(paste("pig ", log[2](FC)))) +
  scale_y_continuous(breaks = seq(-4, 9, by = 1), limits = c(-4, 9)) +
  scale_x_continuous(breaks = seq(-4, 9, by = 1), limits = c(-4, 9))
corr_plot_adjPval_1
```


# AdjPval 0.05 Find shared and unique DEGs between mouse and pig
```{r}
# rename GeneName to gene_name to merge data sets and find intersections
mouse_and_pg_df <-
  merge(mouse_limma_FDRq1.00_orth_1n, pig_brain_FDRq1.00_orth_1n, by = "ortholog_name")

mouse_limma_FDRq0.05_orth_1n <- subset(mouse_limma_FDRq1.00_orth_1n, adj.P.Val < 0.05)
pig_brain_FDRq0.05_orth_1n <- subset(pig_brain_FDRq1.00_orth_1n, adj.P.Val < 0.05)

shared_gene_name <-
  intersect(mouse_limma_FDRq0.05_orth_1n$ortholog_name,
            pig_brain_FDRq0.05_orth_1n$ortholog_name)
uniqueToMouse_gene_name <-
  setdiff(mouse_limma_FDRq0.05_orth_1n$ortholog_name,
          pig_brain_FDRq0.05_orth_1n$ortholog_name)
uniqueToPig_gene_name <-
  setdiff(pig_brain_FDRq0.05_orth_1n$ortholog_name,
          mouse_limma_FDRq0.05_orth_1n$ortholog_name)

shared <-
  mouse_and_pg_df[mouse_and_pg_df$ortholog_name %in% shared_gene_name,]
uniqueToMouse <-
  mouse_and_pg_df[mouse_and_pg_df$ortholog_name %in% uniqueToMouse_gene_name,]
uniqueToPig <-
  mouse_and_pg_df[mouse_and_pg_df$ortholog_name %in% uniqueToPig_gene_name,]


# Venn diagram to see total DEGs between mouse and pig
x = list(mouse = mouse_limma_FDRq0.05_orth_1n$ortholog_name,
         pig = pig_brain_FDRq0.05_orth_1n$ortholog_name)

mouse_pig_venn <- ggvenn(
  x,
  fill_color = c("#EFC000FF", "purple"),
  stroke_size = 2,
  set_name_size = 6
)
mouse_pig_venn
```
## correlation plot - adj.pval < 0.05
```{r}
corr_plot_adjPval_05 <- ggplot(data = shared, aes(x = logFC.x, y = logFC.y, text = paste(gene_name))) +
   annotate(
    "rect",
    xmin = 0,
    xmax = 7,
    ymin = 7,
    ymax = 0,
    fill = "lightpink3",
    alpha = .5
  ) +  annotate(
    "rect",
    xmin = 0,
    xmax = -4,
    ymin = 0,
    ymax = -4,
    fill = "cadetblue3",
    alpha = .5
  ) +
  geom_abline(color = "gray40") +
  geom_point(size = 1) + 
  theme_bw() +
  theme(plot.title = element_text(size = 7, margin = margin(0,0,0,0))) +
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size = 6),
        axis.text.x = element_text(size = 6)) +
  theme(axis.title.y = element_text(size = 6),
        axis.text.y = element_text(size = 6)) +
  labs(
  title = "Shared DEGs between mouse and pig brains
adjusted p.value < 0.05",
  x = expression(paste("mouse ", log[2](FC))),
  y = expression(paste("pig ", log[2](FC)))) +
  scale_y_continuous(breaks = seq(-4, 7, by = 1), limits = c(-4, 7)) +
  scale_x_continuous(breaks = seq(-4, 7, by = 1), limits = c(-4, 7))
corr_plot_adjPval_05
```

# Upset plot
```{r}
mouse_down <- subset(mouse_limma_FDRq0.05_orth_1n, logFC < 0)
mouse_up <- subset(mouse_limma_FDRq0.05_orth_1n, logFC > 0)
pig_down <- subset(pig_brain_FDRq0.05_orth_1n, logFC < 0)
pig_up <- subset(pig_brain_FDRq0.05_orth_1n, logFC > 0)

mouse_down <- mouse_down$ortholog_name
mouse_up <- mouse_up$ortholog_name
pig_down <- pig_down$ortholog_name
pig_up <- pig_up$ortholog_name

list_input <- list("mouse down" = mouse_down,
                   "pig down" = pig_down,
                   "pig up" = pig_up,
                  "mouse up" = mouse_up)
data <- fromList(list_input)

upset_gene <- upset(data, 
      c('mouse down','pig down',
        'mouse up', 'pig up'),
      themes=upset_modify_themes(
        list('intersections_matrix'=theme(text=element_text(size=6)),
          'overall_sizes' = theme(axis.text.x = element_text(size =6)))),
  queries=list(
    upset_query(set='mouse down', fill='blue'),
    upset_query(set='pig down', fill='blue'),
    upset_query(set='mouse up', fill='red'),
    upset_query(set='pig up', fill='red'),
    upset_query(intersect = c('pig down','mouse down'), color='blue',fill='blue', only_components=c('intersections_matrix', 'Intersection size')),
    upset_query(intersect = c('pig up','mouse up'), color='red',fill='red', only_components=c('intersections_matrix', 'Intersection size'))
  ),
  intersections =list(c('mouse down', 'pig down'), 
                      'pig down', 'mouse down', 
                      c('mouse up', 'pig up'), 
                      'pig up', 'mouse up',
                      c('mouse down', 'pig up'), 
                      c('mouse up', 'pig down')
                      ),
  
  base_annotations=list(
    'Intersection size'=(
        intersection_size(
        size=1,
        text = list(size = 2),
        text_mapping = aes(),
        bar_number_threshold=1,  # show all numbers on top of bars
        width=0.3,   # reduce width of the bars
        mapping=aes(fill='bars_color')
      )
      + scale_fill_manual(values=c('grey'), guide='none')
      + scale_y_continuous(expand=expansion(mult=c(0, 0.1)))
      + theme(axis.text.y = element_text(size = 6),
              axis.title.y = element_text(size = 6),
          panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),
          axis.line=element_line(colour='black')))),
  matrix=intersection_matrix(
      geom=geom_point(
        shape='circle filled',
        size=2,
        stroke=0.45)) +
        theme(axis.text.y = element_text(size = 6),
              axis.title.y = element_text(size = 6)),
  set_sizes=(
    upset_set_size(geom=geom_bar(width=0.4), position="right")
    + theme(axis.line.x=element_line(colour='black'), 
            axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
            axis.title.x = element_text(size = 6),
            axis.ticks.x=element_line())),
  sort_sets=FALSE,
  sort_intersections=FALSE)
upset_gene
```
# Metascape Enrichment 
### Format the dataframes
```{r}
# read in enrichment analysis results
up_enrich_results <-
  read.delim(
    paste0(
      "../../../results/",
      tool,
      "/metascape/shared_DEGs_up.txt"
    ),
    sep = "\t",
    header = TRUE
  )

down_enrich_results <-
  read.delim(
    paste0(
      "../../../results/",
      tool,
      "/metascape/shared_DEGs_down.txt"
    ),
    sep = "\t",
    header = TRUE
  )

# select the GO term IDs we want to show in the plot
GO_ID_up <-
  c(
    "1_Summary",
    "2_Summary",
    "3_Summary",
    "4_Summary",
    "5_Summary",
    "6_Summary",
    "7_Summary",
    "8_Summary",
    "9_Summary", 
    "10_Summary",
    "11_Summary",
    "12_Summary",
    "13_Summary",
    "14_Summary",
    "15_Summary",
    "16_Summary",
    "17_Summary",
    "18_Summary",
    "19_Summary", 
    "20_Summary"
  )
GO_ID_down <- c(
    "1_Summary"
)

up_enrich_results_subset <-
  up_enrich_results[up_enrich_results$GroupID %in% GO_ID_up, ]
up_enrich_results_subset$Cluster <- c("up-regulated")
up_enrich_results_subset$Description <-
  factor(up_enrich_results_subset$Description,
         levels = up_enrich_results_subset$Description)
up_enrich_results_subset$Description <-
  fct_rev(up_enrich_results_subset$Description)

down_enrich_results_subset <-
  down_enrich_results[down_enrich_results$GroupID %in% GO_ID_down, ]
down_enrich_results_subset$Cluster <- c("down-regulated")
down_enrich_results_subset$Description <-
  factor(down_enrich_results_subset$Description,
         levels = down_enrich_results_subset$Description)
down_enrich_results_subset$Description <-
  fct_rev(down_enrich_results_subset$Description)

# get the number of genes in each summary
up_gene_count <-
  strsplit(as.character(up_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
up_gene_count_df <-
  data.frame(matrix(
    unlist(up_gene_count),
    nrow = length(up_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
up_enrich_results_subset$InTerm <- as.numeric(up_gene_count_df$X1)
up_enrich_results_subset$InList <- as.numeric(up_gene_count_df$X2)

down_gene_count <-
  strsplit(as.character(down_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
down_gene_count_df <-
  data.frame(matrix(
    unlist(down_gene_count),
    nrow = length(down_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
down_enrich_results_subset$InTerm <- as.numeric(down_gene_count_df$X1)
down_enrich_results_subset$InList <- as.numeric(down_gene_count_df$X2)

# combine together
up_and_down_enrich_results_subset <- rbind(up_enrich_results_subset, down_enrich_results_subset)

#up_and_down_enrich_results_subset$Description <-
 #     gsub(
 #       "regulation of plasma membrane bounded cell projection organization",
#        "reg of plasma membrane/cell projection organization",
 #       up_and_down_enrich_results_subset$Description
  #    )
up_and_down_enrich_results_subset$Description <-
  factor(up_and_down_enrich_results_subset$Description, levels = unique(up_and_down_enrich_results_subset$Description)) 

up_and_down_enrich_results_subset$Description <- fct_rev(up_and_down_enrich_results_subset$Description)
```

remove files 
```{r}
remove(
  down_enrich_results,
  down_gene_count,
  down_gene_count_df,
  up_enrich_results,
  up_gene_count,
  up_gene_count_df
)
```

### Enrichment plot showing log10 P-value
#### down and up seperately
```{r}
down_enrich_plot <-
  ggplot(data = down_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster) +
  geom_bar(stat = "identity", aes(fill = LogP), width = .7, position = position_dodge(width = .2)) +
 ggtitle("shared down DEGs enrichment summaries") +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = "Gene count", y = "") +
  #  xlim(0,60) +
  guides(fill = guide_legend(
    size = 1,
    title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)")
  )) +
  scale_fill_gradientn(
    colours = c("darkmagenta", "mediumorchid3", "grey"),
    guide = "legend",
    limits = c(-45,0),
  ) +
    theme(strip.text = element_text(size = 6), 
          axis.text.y = element_text(size = 7),
          axis.text.x = element_text(size = 6), 
          axis.title.x = element_text(size = 6), 
          legend.text = element_text(size = 6),
          legend.title = element_text(size = 6),
          legend.margin=margin(0,0.5,0,0),
          legend.box.margin=margin(-10,-2,-10,-7.5), 
          plot.margin = margin(0.1, 0.2, 0, 0.2, "cm"), 
          plot.title = element_text(size = 7, hjust = -3.25, vjust=0, margin = margin(0,0,0,0)))
down_enrich_plot

up_enrich_plot <-
  ggplot(data = up_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster) +
  geom_bar(stat = "identity", aes(fill = LogP), width = .7, position = position_dodge(width = .2)) +
  ggtitle("shared up DEGs enrichment summaries") +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = "Gene count", y = "") +
  guides(fill = guide_legend(title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)"))) +
  scale_fill_gradientn(
    colours = c("darkmagenta", "mediumorchid3", "grey"),
    guide = "legend",
    limits = c(-45,-1)
  ) +
    theme(strip.text = element_text(size = 6), 
          axis.text.y = element_text(size = 7),
          axis.text.x = element_text(size = 6), 
          axis.title.x = element_text(size = 6), 
          legend.text = element_text(size = 6),
          legend.title = element_text(size = 6),
          legend.margin=margin(0,0.5,0,0),
          legend.box.margin=margin(-10,-2,-10,-7.5), 
          plot.margin = margin(0.1, 0.2, 0, 0.2, "cm"), 
          plot.title = element_text(size = 7, hjust = -3.25, vjust=0, margin = margin(0,0,0,0)))
up_enrich_plot
```

#### down and up together
```{r}
up_and_down_enrich_plot <-
  ggplot(data = up_and_down_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster, scales = "free") +
  geom_bar(stat = "identity", aes(fill = LogP), width = .7, position = position_dodge(width = .2)) +
  theme_bw() +
  ggtitle(paste0("Shared DEGs enrichment summaries")) +
  labs(x = "Gene count", y = NULL) +
  guides(fill = guide_legend(title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)"))) +
  scale_fill_gradientn(
    colours = c("darkmagenta", "mediumorchid3", "grey"),
    guide = "legend",
    limits = c(-45,-1)
  ) +
    theme(strip.text = element_text(size = 6), 
          axis.text.y = element_text(size = 7),
          axis.text.x = element_text(size = 6), 
          axis.title.x = element_text(size = 6), 
          legend.text = element_text(size = 6),
          legend.title = element_text(size = 6),
          legend.margin=margin(0,0.5,0,0),
          legend.box.margin=margin(-10,-2,-10,-7.5), 
          plot.margin = margin(0.1, 0.2, 0, 0.2, "cm"), 
          plot.title = element_text(size = 7, hjust = -3.25, vjust=0, margin = margin(0,0,1,0)))
up_and_down_enrich_plot

addSmallLegend <- function(myPlot, pointSize = 3, textSize = 6, spaceLegend = .5) {
    myPlot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}

# Apply on original plot
up_and_down_enrich_plot <- addSmallLegend(up_and_down_enrich_plot)
```

# Heatmaps
```{r}
mouse_down <- subset(mouse_limma_FDRq0.05_orth_1n, logFC < 0)
mouse_up <- subset(mouse_limma_FDRq0.05_orth_1n, logFC > 0)
pig_down <- subset(pig_brain_FDRq0.05_orth_1n, logFC < 0)
pig_up <- subset(pig_brain_FDRq0.05_orth_1n, logFC > 0)


pig_up_mouse_down_ortholog_name <- intersect(pig_up$ortholog_name, mouse_down$ortholog_name)
pig_down_mouse_up_ortholog_name <- intersect(pig_down$ortholog_name, mouse_up$ortholog_name)
write.table(pig_down_mouse_up_ortholog_name, 
            "../../../mouse_comparison/results/pig_down_mouse_up_ortholog_name.txt", 
            sep = "\t", 
            quote = FALSE, 
            row.names = FALSE)
write.table(pig_up_mouse_down_ortholog_name, 
            "../../../mouse_comparison/results/pig_up_mouse_down_ortholog_name.txt", 
            sep = "\t", 
            quote = FALSE, 
            row.names = FALSE)

pig_up$species <- c("pig")
mouse_down$species <- c("mouse")
pig_down$species <- c("pig")
mouse_up$species <- c("mouse")

df_pig_up_mouse_down <- smartbind(pig_up, mouse_down)
df_pig_down_mouse_up <- smartbind(pig_down, mouse_up)

# get the fold change value for those genes
df_pig_up_mouse_down_sub <- subset(df_pig_up_mouse_down, ortholog_name %in% pig_up_mouse_down_ortholog_name)
df_pig_down_mouse_up_sub <- subset(df_pig_down_mouse_up, ortholog_name %in% pig_down_mouse_up_ortholog_name)

# order by fold change 
df_pig_up_mouse_down_sub <- df_pig_up_mouse_down_sub[order(-df_pig_up_mouse_down_sub$logFC),]
df_pig_up_mouse_down_sub$species <- factor(df_pig_up_mouse_down_sub$species, levels = c("pig", "mouse"))

df_pig_down_mouse_up_sub <- df_pig_down_mouse_up_sub[order(df_pig_down_mouse_up_sub$logFC),]
df_pig_down_mouse_up_sub$species <- factor(df_pig_down_mouse_up_sub$species, levels = c("pig", "mouse"))
```
## up in pigs & down in mouse
```{r}
# lock in gene order which is sorted by adjusted p-value
df_pig_up_mouse_down_sub$ortholog_name <- factor(df_pig_up_mouse_down_sub$ortholog_name,
                          levels = unique(df_pig_up_mouse_down_sub$ortholog_name))
df_pig_up_mouse_down_sub$ortholog_name<- fct_rev(df_pig_up_mouse_down_sub$ortholog_name)
# heatmap 
heatmap1 <- ggplot(data = df_pig_up_mouse_down_sub, aes(x = species, y = ortholog_name)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ species, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "navyblue",
    mid = "white",
    high = "maroon",
    midpoint = 0,
    space = "rgb",
    guide = "colourbar",
  #  breaks = c(-2.5, 0, 4.5),
   # limits = c(-2.5,4.5),
    name = expression(log[2](FC))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
#    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0.2, 0.5, 0.2, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("up-regulated in pigs
down-regulated in mouse") 
heatmap1

# Apply on original plot
heatmap1 <- addSmallLegend(heatmap1)
```
## down in pigs & up in mouse
```{r}
# lock in gene order which is sorted by adjusted p-value
df_pig_down_mouse_up_sub$ortholog_name <- factor(df_pig_down_mouse_up_sub$ortholog_name,
                          levels = unique(df_pig_down_mouse_up_sub$ortholog_name))
df_pig_down_mouse_up_sub$ortholog_name<- fct_rev(df_pig_down_mouse_up_sub$ortholog_name)
# heatmap 
heatmap2 <- ggplot(data = df_pig_down_mouse_up_sub, aes(x = species, y = ortholog_name)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ species, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "navyblue",
    mid = "white",
    high = "maroon",
    midpoint = 0,
    space = "rgb",
    guide = "colourbar",
  #  breaks = c(-3.5, 0, 2.5),
  #  limits = c(-3.5,2.5),
    name = expression(log[2](FC))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
   # legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0.2, 0.5, 0.2, 0.7, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("down-regulated in pigs
up-regulated in mouse") 
heatmap2

heatmap2 <- addSmallLegend(heatmap2)
```

# Combine plot 
```{r}
col1 <- ggarrange(
  upset_gene,
  corr_plot_adjPval_1,
  corr_plot_adjPval_05,
  ncol = 1, 
  nrow = 3,
  labels = c("a)", "b)", "c)"),
  font.label = list(size = 8)
  )

col2 <-
  ggarrange( 
    heatmap1,
    heatmap2,
    nrow = 2,
    labels = c("d)", "e)"),
    heights = c(1.65, 1),
    font.label = list(size = 8)
  )

together <- ggarrange(col1, col2, ncol = 2, widths = c(1.5, 1))
together
path <- paste0("../../../results/manuscript_figures/Figure_",
               "mouse_pig",
               "_correlation_v2")
saveToPDF(paste0(path, ".pdf"), width = 7.08, height = 8.66)
```
# session info
```{r}
```