---
title: "Upset Plot"
author: "Kimberly Olney"
date: "6/14/2022"
output: html_document
---

# Setup
```{r setup, include=FALSE}
# Also do Session > Set Working Directory > Choose Directory
knitr::opts_knit$set(root.dir = ".")
```

```{r packages}
library(UpSetR)
library(ComplexUpset)
library(ggplot2movies)
library(ggpubr)
library(forcats)
library(stringr)
```

```{r}
saveToPDF <- function(...) {
    d = dev.copy(pdf,...)
    dev.off(d)
}
```
# Binary list shared gene function
```{r}
fromList <- function (input) {
  # Same as original fromList()...
  elements <- unique(unlist(input))
  data <- unlist(lapply(input, function(x) {
      x <- as.vector(match(elements, x))
      }))
  data[is.na(data)] <- as.integer(0)
  data[data != 0] <- as.integer(1)
  data <- data.frame(matrix(data, ncol = length(input), byrow = F))
  data <- data[which(rowSums(data) != 0), ]
  names(data) <- names(input)
  # ... Except now it conserves your original value names!
  row.names(data) <- elements
  return(data)
  }
```

# User defined variables
```{r set_variables}
tissue <- "shared_bt_brain_and_kidney"
control <- "Saline"
treatment <- "LPS"
control_color <- "gray29"
treatment_color <- "purple"
tool = c("star")
```

# Read in DEGs
```{r}
star_brain <- read.delim("../../../results/star/DEGs/LPS_Brain_gene_DEGs_FDRq0.05.txt", 
                    header = TRUE, sep = "\t")
star_kidney <- read.delim("../../../results/star/DEGs/LPS_Kidney_gene_DEGs_FDRq0.05.txt", 
                     header = TRUE, sep = "\t")
```

# Subset lists
```{r}
up_star_brain <- subset(star_brain$gene_name, star_brain$logFC > 0)
down_star_brain <- subset(star_brain$gene_name, star_brain$logFC < 0)

up_star_kidney <- subset(star_kidney$gene_name, star_kidney$logFC > 0)
down_star_kidney <- subset(star_kidney$gene_name, star_kidney$logFC < 0)
```

# Upset plot between kidney and brain
```{r}
list_input <- list("Brain down-regulated" = down_star_brain,
                   "Brain up-regulated" = up_star_brain,
                   "Kidney down-regulated" = down_star_kidney,
                   "Kidney up-regulated" = up_star_kidney)
data <- fromList(list_input)

upset_gene <- upset(data, set_sizes=FALSE,
      c('Brain down-regulated','Brain up-regulated','Kidney down-regulated', 
      'Kidney up-regulated'),
  queries=list(
    upset_query(set='Brain down-regulated', fill='blue'),
    upset_query(set='Brain up-regulated', fill='red'),
    upset_query(set='Kidney down-regulated', fill='blue'),
    upset_query(set='Kidney up-regulated', fill='red')
  ),
  base_annotations=list(
    'Intersection size'=(
        intersection_size(
        bar_number_threshold=1,  # show all numbers on top of bars
        width=0.5,   # reduce width of the bars
      )
      # add some space on the top of the bars
      + scale_y_continuous(expand=expansion(mult=c(0, 0.05)))
      + theme(
          # hide grid lines
          panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),
          # show axis lines
          axis.line=element_line(colour='black')
      )
    )
  ),
  stripes=upset_stripes(
    geom=geom_segment(size=12),  # make the stripes larger
    colors=c('grey95', 'white')
  ),
  # to prevent connectors from getting the colorured
  # use `fill` instead of `color`, together with `shape='circle filled'`
  matrix=intersection_matrix(
      geom=geom_point(
        shape='circle filled',
        size=3,
        stroke=0.45
      )
  ),
  sort_sets='descending',
  sort_intersections='descending'
)
upset_gene
```

# Upset plot pretty
```{r}
list_input <- list("Brain down" = down_star_brain,
                   "Kidney down" = down_star_kidney,
                   "Kidney up" = up_star_kidney,
                  "Brain up" = up_star_brain)
data <- fromList(list_input)

upset_gene <- upset(data, 
      c('Brain down','Kidney down',
        'Brain up', 'Kidney up'),
      themes=upset_modify_themes(
        list('intersections_matrix'=theme(text=element_text(size=6)),
          'overall_sizes' = theme(axis.text.x = element_text(size =6)))),
  queries=list(
    upset_query(set='Brain down', fill='blue'),
    upset_query(set='Kidney down', fill='blue'),
    upset_query(set='Brain up', fill='red'),
    upset_query(set='Kidney up', fill='red'),
    upset_query(intersect = c('Kidney down','Brain down'), color='blue',fill='blue', only_components=c('intersections_matrix', 'Intersection size')),
    upset_query(intersect = c('Kidney up','Brain up'), color='red',fill='red', only_components=c('intersections_matrix', 'Intersection size'))
  ),
  intersections =list(c('Brain down', 'Kidney down'), 
                      'Kidney down', 'Brain down', 
                      c('Brain up', 'Kidney up'), 
                      'Kidney up', 'Brain up',
                      c('Brain down', 'Kidney up'), 
                      c('Brain up', 'Kidney down')
                      ),
  
  base_annotations=list(
    'Intersection size'=(
        intersection_size(
        size=1,
        text = list(size = 2),
        text_mapping = aes(),
        bar_number_threshold=1,  # show all numbers on top of bars
        width=0.3,   # reduce width of the bars
        mapping=aes(fill='bars_color')
      )
      + scale_fill_manual(values=c('grey'), guide='none')
      + scale_y_continuous(expand=expansion(mult=c(0, 0.1)))
      + theme(axis.text.y = element_text(size = 6),
              axis.title.y = element_text(size = 6),
          panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),
          axis.line=element_line(colour='black')))),
  matrix=intersection_matrix(
      geom=geom_point(
        shape='circle filled',
        size=2,
        stroke=0.45)) +
        theme(axis.text.y = element_text(size = 6),
              axis.title.y = element_text(size = 6)),
  set_sizes=(
    upset_set_size(geom=geom_bar(width=0.4), position="right")
    + theme(axis.line.x=element_line(colour='black'), 
            axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
            axis.title.x = element_text(size = 6),
            axis.ticks.x=element_line())),
  sort_sets=FALSE,
  sort_intersections=FALSE)
upset_gene
```
# Correlation plot
## read in pig data 
```{r}
# brain
pig_brain_FDRq0.05 <-
  read.delim("../../../results/star/DEGs/LPS_Brain_gene_DEGs_FDRq0.05.txt")
pig_brain_FDRq1.00 <-
  read.delim("../../../results/star/DEGs/LPS_Brain_gene_DEGs_FDRq1.00.txt")

# kidney
pig_kidney_FDRq0.05 <-
  read.delim("../../../results/star/DEGs/LPS_Kidney_gene_DEGs_FDRq0.05.txt")
pig_kidney_FDRq1.00 <-
  read.delim("../../../results/star/DEGs/LPS_Kidney_gene_DEGs_FDRq1.00.txt")
```

## shared and unique between brain and kidney 
```{r}
# rename GeneName to gene_name to merge data sets and find intersections
kidneyBrain <- merge(pig_kidney_FDRq1.00, pig_brain_FDRq1.00, by = "gene_name")

shared_gene_name <- intersect(pig_brain_FDRq0.05$gene_name, pig_kidney_FDRq0.05$gene_name)
uniqueToKidney_gene_name <- setdiff(pig_kidney_FDRq0.05$gene_name, pig_brain_FDRq0.05$gene_name)
uniqueToBrain_gene_name <- setdiff(pig_brain_FDRq0.05$gene_name, pig_kidney_FDRq0.05$gene_name)

shared <- kidneyBrain[kidneyBrain$gene_name %in% shared_gene_name, ]
uniqueToBrain <- kidneyBrain[kidneyBrain$gene_name %in% uniqueToBrain_gene_name, ]
uniqueToKidney <- kidneyBrain[kidneyBrain$gene_name %in% uniqueToKidney_gene_name, ]
# Venn diagram to see what is gained and lost between the reanalysis
x = list(brain = pig_brain_FDRq0.05$gene_name, kidney = pig_kidney_FDRq0.05$gene_name)
ggvenn(
  x,
  fill_color = c("#EFC000FF", "blue"),
  stroke_size = 2,
  set_name_size = 6
)
```

## correlation plot - all
```{r}
# Shapiro-Wilk normality test 
shapiro.test(shared$logFC.x)
shapiro.test(shared$logFC.y)

# correlation test 
res <- cor.test(shared$logFC.x, shared$logFC.y, 
                    method = "spearman")
p_value <- round(res$p.value, 3)
rho_value <- round(res$estimate, 3) 

corr_plot_shared_DEGs <- ggplot(data = shared, aes(x = logFC.x, y = logFC.y, text = paste(gene_name))) +
   annotate(
    "rect",
    xmin = 0,
    xmax = 9,
    ymin = 9,
    ymax = 0,
    fill = "lightpink3",
    alpha = .5
  ) +  annotate(
    "rect",
    xmin = 0,
    xmax = -6,
    ymin = 0,
    ymax = -6,
    fill = "cadetblue3",
    alpha = .5
  ) +
  geom_abline(color = "gray40") +
  geom_point(size = 1) + 
  theme_bw() +
  theme(plot.title = element_text(size = 7, margin = margin(0,0,0,0))) +
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size = 6),
        axis.text.x = element_text(size = 6)) +
  theme(axis.title.y = element_text(size = 6),
        axis.text.y = element_text(size = 6)) +
  labs(
  title = "Shared DEGs between brain and kidney",
  x = expression(paste("kidney ", log[2](FC))),
  y = expression(paste("brain ", log[2](FC)))) +
  scale_y_continuous(breaks = seq(-6, 9, by = 1), limits = c(-6, 9)) +
  scale_x_continuous(breaks = seq(-6, 9, by = 1), limits = c(-6, 9))
corr_plot_shared_DEGs
```
# Metascape Enrichment 
### Format the dataframes
```{r}
# read in enrichment analysis results
up_enrich_results <-
  read.delim(
    paste0(
      "../../../results/",
      tool,
      "/metascape/shared_DEGs_brain_kidney_up.txt"
    ),
    sep = "\t",
    header = TRUE
  )

down_enrich_results <-
  read.delim(
    paste0(
      "../../../results/",
      tool,
      "/metascape/shared_DEGs_brain_kidney_down.txt"
    ),
    sep = "\t",
    header = TRUE
  )

# select the GO term IDs we want to show in the plot
GO_ID_up <-
  c(
    "1_Summary",
    "2_Summary",
    "3_Summary",
    "4_Summary",
    "5_Summary",
    "6_Summary",
    "7_Summary",
    "8_Summary",
    "9_Summary", 
    "10_Summary",
    "11_Summary",
    "12_Summary",
    "13_Summary",
    "14_Summary",
    "15_Summary",
    "16_Summary",
    "17_Summary",
    "18_Summary",
    "19_Summary", 
    "20_Summary"
  )
GO_ID_down <- c(
    "1_Summary",
    "2_Summary",
    "3_Summary",
    "4_Summary",
    "5_Summary",
    "6_Summary",
    "7_Summary",
    "8_Summary",
    "9_Summary", 
    "10_Summary",
    "11_Summary"
)

up_enrich_results_subset <-
  up_enrich_results[up_enrich_results$GroupID %in% GO_ID_up, ]
up_enrich_results_subset$Cluster <- c("up-regulated")
up_enrich_results_subset$Description <-
  factor(up_enrich_results_subset$Description,
         levels = up_enrich_results_subset$Description)
up_enrich_results_subset$Description <-
  fct_rev(up_enrich_results_subset$Description)

down_enrich_results_subset <-
  down_enrich_results[down_enrich_results$GroupID %in% GO_ID_down, ]
down_enrich_results_subset$Cluster <- c("down-regulated")
down_enrich_results_subset$Description <-
  factor(down_enrich_results_subset$Description,
         levels = down_enrich_results_subset$Description)
down_enrich_results_subset$Description <-
  fct_rev(down_enrich_results_subset$Description)

# get the number of genes in each summary
up_gene_count <-
  strsplit(as.character(up_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
up_gene_count_df <-
  data.frame(matrix(
    unlist(up_gene_count),
    nrow = length(up_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
up_enrich_results_subset$InTerm <- as.numeric(up_gene_count_df$X1)
up_enrich_results_subset$InList <- as.numeric(up_gene_count_df$X2)

down_gene_count <-
  strsplit(as.character(down_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
down_gene_count_df <-
  data.frame(matrix(
    unlist(down_gene_count),
    nrow = length(down_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
down_enrich_results_subset$InTerm <- as.numeric(down_gene_count_df$X1)
down_enrich_results_subset$InList <- as.numeric(down_gene_count_df$X2)

# combine together
up_and_down_enrich_results_subset <- rbind(up_enrich_results_subset, down_enrich_results_subset)

up_and_down_enrich_results_subset$Description <- gsub(
  "TNF-related weak inducer of apoptosis (TWEAK) signaling pathway",
  "TNF weak inducer of apoptosis signaling pathway", fixed=TRUE,
  up_and_down_enrich_results_subset$Description)


up_and_down_enrich_results_subset$Description <-
  factor(up_and_down_enrich_results_subset$Description, levels = unique(up_and_down_enrich_results_subset$Description)) 

up_and_down_enrich_results_subset$Description <- fct_rev(up_and_down_enrich_results_subset$Description)
```

remove files 
```{r}
remove(
  down_enrich_results,
  down_gene_count,
  down_gene_count_df,
  up_enrich_results,
  up_gene_count,
  up_gene_count_df
)
```
### Genes in enrichment plot all
## up 
[1] Cytokine Signaling in Immune system                            
 [2] response to virus                                              
 [3] TNF signaling pathway                                          
 [4] response to lipopolysaccharide                                 
 [5] regulation of defense response                                 
 [6] positive regulation of cytokine production                     
 [7] Interferon Signaling                                           
 [8] positive regulation of cell death                              
 [9] Epstein-Barr virus infection                                   
[10] regulation of response to biotic stimulus                      
[11] TNF-related weak inducer of apoptosis (TWEAK) signaling pathway
[12] PID P53 DOWNSTREAM PATHWAY                                     
[13] Orexin receptor pathway                                        
[14] immune system development                                      
[15] Lipid and atherosclerosis                                      
[16] VEGFA-VEGFR2 signaling pathway                                 
[17] Human T-cell leukemia virus 1 infection                        
[18] response to tumor necrosis factor                              
[19] cellular response to external stimulus                         
[20] regulation of response to cytokine stimulus 
```{r}
up_enrich_df <-
  up_enrich_results_subset[, c("GroupID", "Term", "Description", "Symbols")]
up_enrich_df_melt <- reshape2::melt(up_enrich_df)
up_gene_char <- str_split(up_enrich_df_melt$Symbols, ",")

up_gene_pathway <-
  data.frame(gene = unlist(up_gene_char),
             value = rep(up_enrich_df$Description, lengths(up_gene_char)))
up_pathways_levels <- levels(fct_rev(up_gene_pathway$value))

up_input <- list(
  "Cytokine Signaling in Immune system" = up_gene_char[[1]],
  "Interferon Signaling" = up_gene_char[[7]],
  "positive regulation of cell death" = up_gene_char[[8]],
  "VEGFA-VEGFR2 signaling pathway" = up_gene_char[[16]]
)
up_data <- fromList(up_input)
up_data$gene <- row.names(up_data)
up_data_melt <- reshape2::melt(up_data)

# get the fold change value for those genes
shared$gene <- shared$gene_name
up_df <- merge(up_data_melt, shared,
                       by = "gene")

# subset by GO termes
GO_1 <-
  subset(up_df,
         up_df$variable == "Cytokine Signaling in Immune system"
         & up_df$value == 1)
GO_2 <-
  subset(up_df,
         up_df$variable == "Interferon Signaling"
         & up_df$value == 1)
GO_3 <-
  subset(up_df,
         up_df$variable == "positive regulation of cell death"
         & up_df$value == 1)
GO_4 <-
  subset(up_df,
         up_df$variable == "VEGFA-VEGFR2 signaling pathway"
         & up_df$value == 1)
GO_1$gene
```
### up 1 correlation
```{r}
# Shapiro-Wilk normality test 
shapiro.test(shared$logFC.x)
shapiro.test(shared$logFC.y)

# correlation test 
res <- cor.test(shared$logFC.x, shared$logFC.y, 
                    method = "spearman")
p_value <- round(res$p.value, 3)
rho_value <- round(res$estimate, 3) 

up1_corr_plot <- ggplot(data = GO_1, aes(x = logFC.x, y = logFC.y, text = paste(gene_name))) +
   annotate(
    "rect",
    xmin = 0,
    xmax = 9,
    ymin = 9,
    ymax = 0,
    fill = "lightpink3",
    alpha = .5
  ) +  
  geom_abline(color = "gray40") +
  geom_point(size = 1) + 
  theme_bw() +
  theme(plot.title = element_text(size = 7,margin = margin(0,0,0,0))) +
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size = 6),
        axis.text.x = element_text(size = 6)) +
  theme(axis.title.y = element_text(size = 6),
        axis.text.y = element_text(size = 6)) +
  labs(
  title = "Cytokine Signaling 
in Immune system",
  x = expression(paste("kidney ", log[2](FC))),
  y = expression(paste("brain ", log[2](FC)))) +
  scale_y_continuous(breaks = seq(0, 9, by = 1), limits = c(0, 9)) +
  scale_x_continuous(breaks = seq(0, 9, by = 1), limits = c(0, 9))
up1_corr_plot
```
### up 2 correlation
```{r}
# Shapiro-Wilk normality test 
shapiro.test(shared$logFC.x)
shapiro.test(shared$logFC.y)

# correlation test 
res <- cor.test(shared$logFC.x, shared$logFC.y, 
                    method = "spearman")
p_value <- round(res$p.value, 3)
rho_value <- round(res$estimate, 3) 

up2_corr_plot <- ggplot(data = GO_2, aes(x = logFC.x, y = logFC.y, text = paste(gene_name))) +
   annotate(
    "rect",
    xmin = 0,
    xmax = 9,
    ymin = 9,
    ymax = 0,
    fill = "lightpink3",
    alpha = .5
  ) +  
  geom_abline(color = "gray40") +
  geom_point(size = 1) + 
  theme_bw() +
  theme(plot.title = element_text(size = 7,margin = margin(0,0,0,0))) +
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size = 6),
        axis.text.x = element_text(size = 6)) +
  theme(axis.title.y = element_text(size = 6),
        axis.text.y = element_text(size = 6)) +
  labs(
  title = "Interferon 
Signaling",
  x = expression(paste("kidney ", log[2](FC))),
  y = expression(paste("brain ", log[2](FC)))) +
  scale_y_continuous(breaks = seq(0, 9, by = 1), limits = c(0, 9)) +
  scale_x_continuous(breaks = seq(0, 9, by = 1), limits = c(0, 9))
up2_corr_plot
```

### up 3 correlation
```{r}
# Shapiro-Wilk normality test 
shapiro.test(shared$logFC.x)
shapiro.test(shared$logFC.y)

# correlation test 
res <- cor.test(shared$logFC.x, shared$logFC.y, 
                    method = "spearman")
p_value <- round(res$p.value, 3)
rho_value <- round(res$estimate, 3) 

up3_corr_plot <- ggplot(data = GO_3, aes(x = logFC.x, y = logFC.y, text = paste(gene_name))) +
   annotate(
    "rect",
    xmin = 0,
    xmax = 9,
    ymin = 9,
    ymax = 0,
    fill = "lightpink3",
    alpha = .5
  ) +  
  geom_abline(color = "gray40") +
  geom_point(size = 1) + 
  theme_bw() +
  theme(plot.title = element_text(size = 7,margin = margin(0,0,0,0))) +
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size = 6),
        axis.text.x = element_text(size = 6)) +
  theme(axis.title.y = element_text(size = 6),
        axis.text.y = element_text(size = 6)) +
  labs(
  title = "positive regulation 
of cell death",
  x = expression(paste("kidney ", log[2](FC))),
  y = expression(paste("brain ", log[2](FC)))) +
  scale_y_continuous(breaks = seq(0, 9, by = 1), limits = c(0, 9)) +
  scale_x_continuous(breaks = seq(0, 9, by = 1), limits = c(0, 9))
up3_corr_plot
```
### up 4 correlation
```{r}
# Shapiro-Wilk normality test 
shapiro.test(shared$logFC.x)
shapiro.test(shared$logFC.y)

# correlation test 
res <- cor.test(shared$logFC.x, shared$logFC.y, 
                    method = "spearman")
p_value <- round(res$p.value, 3)
rho_value <- round(res$estimate, 3) 

up4_corr_plot <- ggplot(data = GO_4, aes(x = logFC.x, y = logFC.y, text = paste(gene_name))) +
   annotate(
    "rect",
    xmin = 0,
    xmax = 9,
    ymin = 9,
    ymax = 0,
    fill = "lightpink3",
    alpha = .5
  ) +  
  geom_abline(color = "gray40") +
  geom_point(size = 1) + 
  theme_bw() +
  theme(plot.title = element_text(size = 7,margin = margin(0,0,0,0))) +
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size = 6),
        axis.text.x = element_text(size = 6)) +
  theme(axis.title.y = element_text(size = 6),
        axis.text.y = element_text(size = 6)) +
  labs(
  title = "VEGFA-VEGFR2 
signaling pathway",
  x = expression(paste("kidney ", log[2](FC))),
  y = expression(paste("brain ", log[2](FC)))) +
  scale_y_continuous(breaks = seq(0, 9, by = 1), limits = c(0, 9)) +
  scale_x_continuous(breaks = seq(0, 9, by = 1), limits = c(0, 9))
up4_corr_plot
```

### Genes in enrichment plot all
## down 
* [1] angiogenesis                                             
 [2] hematopoietic or lymphoid organ development              
 [3] Hematopoietic stem cell differentiation                  
* [4] lymph vessel development                                 
* [5] endothelium development                                  
 [6] regulation of cytoskeleton organization                  
 [7] Epithelial to mesenchymal transition in colorectal cancer
* [8] RAC1 GTPase cycle                                        
 [9] regulation of protein binding                            
[10] cellular response to growth factor stimulus              
* [11] phagocytosis
```{r}
down_enrich_df <-
  down_enrich_results_subset[, c("GroupID", "Term", "Description", "Symbols")]
down_enrich_df_melt <- reshape2::melt(down_enrich_df)
down_gene_char <- str_split(down_enrich_df_melt$Symbols, ",")

down_gene_pathway <-
  data.frame(gene = unlist(down_gene_char),
             value = rep(down_enrich_df$Description, lengths(down_gene_char)))
down_pathways_levels <- levels(fct_rev(down_gene_pathway$value))

down_input <- list(
  "angiogenesis" = down_gene_char[[1]], 
  "lymph vessel development" = down_gene_char[[4]],
  "endothelium development" = down_gene_char[[5]],
  "RAC1 GTPase cycle" = down_gene_char[[8]]
)
down_data <- fromList(down_input)
down_data$gene <- row.names(down_data)
down_data_melt <- reshape2::melt(down_data)

# get the fold change value for those genes
shared$gene <- shared$gene_name
down_df <- merge(down_data_melt, shared,
                       by = "gene")

# subset by GO termes
GO_1 <-
  subset(down_df,
         down_df$variable == "angiogenesis"
         & down_df$value == 1)
GO_2 <-
  subset(down_df,
         down_df$variable == "lymph vessel development"
         & down_df$value == 1)
GO_3 <-
  subset(down_df,
         down_df$variable == "endothelium development"
         & down_df$value == 1)
GO_4 <-
  subset(down_df,
         down_df$variable == "RAC1 GTPase cycle"
         & down_df$value == 1)
```
### down 1 correlation
```{r}
# Shapiro-Wilk normality test 
shapiro.test(shared$logFC.x)
shapiro.test(shared$logFC.y)

# correlation test 
res <- cor.test(shared$logFC.x, shared$logFC.y, 
                    method = "spearman")
p_value <- round(res$p.value, 3)
rho_value <- round(res$estimate, 3) 

down1_corr_plot <- ggplot(data = GO_1, aes(x = logFC.x, y = logFC.y, text = paste(gene_name))) +
   annotate(
    "rect",
    xmin = 0,
    xmax = -6,
    ymin = -6,
    ymax = 0,
    fill = "cadetblue3",
    alpha = .5
  ) +  
  geom_abline(color = "gray40") +
  geom_point(size = 1) + 
  theme_bw() +
  theme(plot.title = element_text(size = 7,margin = margin(0,0,0,0))) +
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size = 6),
        axis.text.x = element_text(size = 6)) +
  theme(axis.title.y = element_text(size = 6),
        axis.text.y = element_text(size = 6)) +
  labs(
  title = "angiogenesis",
  x = expression(paste("kidney ", log[2](FC))),
  y = expression(paste("brain ", log[2](FC)))) +
  scale_y_continuous(breaks = seq(-6, 0, by = 1), limits = c(-6, 0)) +
  scale_x_continuous(breaks = seq(-6, 0, by = 1), limits = c(-6, 0))
down1_corr_plot
```
### down 2 correlation
```{r}
# Shapiro-Wilk normality test 
shapiro.test(shared$logFC.x)
shapiro.test(shared$logFC.y)

# correlation test 
res <- cor.test(shared$logFC.x, shared$logFC.y, 
                    method = "spearman")
p_value <- round(res$p.value, 3)
rho_value <- round(res$estimate, 3) 

down2_corr_plot <- ggplot(data = GO_2, aes(x = logFC.x, y = logFC.y, text = paste(gene_name))) +
   annotate(
    "rect",
    xmin = 0,
    xmax = -6,
    ymin = -6,
    ymax = 0,
    fill = "cadetblue3",
    alpha = .5
  ) +  
  geom_abline(color = "gray40") +
  geom_point(size = 1) + 
  theme_bw() +
  theme(plot.title = element_text(size = 7,margin = margin(0,0,0,0))) +
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size = 6),
        axis.text.x = element_text(size = 6)) +
  theme(axis.title.y = element_text(size = 6),
        axis.text.y = element_text(size = 6)) +
  labs(
  title = "lymph vessel development",
  x = expression(paste("kidney ", log[2](FC))),
  y = expression(paste("brain ", log[2](FC)))) +
  scale_y_continuous(breaks = seq(-6, 0, by = 1), limits = c(-6, 0)) +
  scale_x_continuous(breaks = seq(-6, 0, by = 1), limits = c(-6, 0))
down2_corr_plot
```

### down 3 correlation
```{r}
# Shapiro-Wilk normality test 
shapiro.test(shared$logFC.x)
shapiro.test(shared$logFC.y)

# correlation test 
res <- cor.test(shared$logFC.x, shared$logFC.y, 
                    method = "spearman")
p_value <- round(res$p.value, 3)
rho_value <- round(res$estimate, 3) 

down3_corr_plot <- ggplot(data = GO_3, aes(x = logFC.x, y = logFC.y, text = paste(gene_name))) +
   annotate(
    "rect",
    xmin = 0,
    xmax = -6,
    ymin = -6,
    ymax = 0,
    fill = "cadetblue3",
    alpha = .5
  ) +  
  geom_abline(color = "gray40") +
  geom_point(size = 1) + 
  theme_bw() +
  theme(plot.title = element_text(size = 7,margin = margin(0,0,0,0))) +
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size = 6),
        axis.text.x = element_text(size = 6)) +
  theme(axis.title.y = element_text(size = 6),
        axis.text.y = element_text(size = 6)) +
  labs(
  title = "endothelium development",
  x = expression(paste("kidney ", log[2](FC))),
  y = expression(paste("brain ", log[2](FC)))) +
  scale_y_continuous(breaks = seq(-6, 0, by = 1), limits = c(-6, 0)) +
  scale_x_continuous(breaks = seq(-6, 0, by = 1), limits = c(-6, 0))
down3_corr_plot
```
### down 4 correlation
```{r}
# Shapiro-Wilk normality test 
shapiro.test(shared$logFC.x)
shapiro.test(shared$logFC.y)

# correlation test 
res <- cor.test(shared$logFC.x, shared$logFC.y, 
                    method = "spearman")
p_value <- round(res$p.value, 3)
rho_value <- round(res$estimate, 3) 

down4_corr_plot <- ggplot(data = GO_4, aes(x = logFC.x, y = logFC.y, text = paste(gene_name))) +
   annotate(
    "rect",
    xmin = 0,
    xmax = -3,
    ymin = -3,
    ymax = 0,
    fill = "cadetblue3",
    alpha = .5
  ) +  
  geom_abline(color = "gray40") +
  geom_point(size = 1) + 
  theme_bw() +
  theme(plot.title = element_text(size = 7,margin = margin(0,0,0,0))) +
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size = 6),
        axis.text.x = element_text(size = 6)) +
  theme(axis.title.y = element_text(size = 6),
        axis.text.y = element_text(size = 6)) +
  labs(
  title = "RAC1 GTPase cycle",
  x = expression(paste("kidney ", log[2](FC))),
  y = expression(paste("brain ", log[2](FC)))) +
  scale_y_continuous(breaks = seq(-3, 0, by = 1), limits = c(-3, 0)) +
  scale_x_continuous(breaks = seq(-3, 0, by = 1), limits = c(-3, 0))
down4_corr_plot
```

# Combine plot 
```{r}
row1_left <- ggarrange(
  upset_gene,
  labels = c("a)"),
  font.label = list(size = 8)
  )
row1_right <- ggarrange(
  corr_plot_shared_DEGs,
  labels = c("b)"),
  font.label = list(size = 8)
)
row1 <- ggarrange(
  row1_left,
  row1_right,
  ncol = 2, 
  widths = c(1, 1)
)
row1


row2 <-
  ggarrange(
    NULL, 
    NULL,
    NULL,
    NULL,
    ncol = 4,
    labels = c("up-regulated", "", "", "", ""),
    font.label = list(size = 7, color = "red")
  )

row3 <- ggarrange(
  up1_corr_plot, 
  up2_corr_plot, 
  up3_corr_plot, 
  up4_corr_plot, 
  ncol = 4,
  labels = c("d)", "e)", "f)", "g)"),
  font.label = list(size = 8)
  )


row4 <-
  ggarrange(
    NULL, 
    NULL,
    NULL,
    NULL,
    ncol = 4,
    labels = c("c)"),
    font.label = list(size = 8)
  )

row5 <-
  ggarrange(
    NULL, 
    NULL,
    NULL,
    NULL,
    ncol = 4,
    labels = c("down-regulated", "", "", "", ""),
    font.label = list(size = 7, color = "blue")
  )

row6 <- ggarrange(
  down1_corr_plot, 
  down2_corr_plot, 
  down3_corr_plot, 
  down4_corr_plot, 
  ncol = 4,
  labels = c("g)", "h)", "i)", "j)"),
  font.label = list(size = 8)
  )
together <- ggarrange(row1, row4, row2, row3, row5, row6, nrow = 6, heights = c(1.25, 1, .075, 1, .075, .75))
together
path <- paste0("../../../results/manuscript_figures/Figure_",
               "shared_DEGs_brain_kidney",
               "_correlation")
saveToPDF(paste0(path, ".pdf"), width = 7.08, height = 8.66)
```
# session info
```{r}
```