---
title: "Upset Plot"
author: "Kimberly Olney"
date: "6/14/2022"
output: html_document
---

# Setup
```{r setup, include=FALSE}
# Also do Session > Set Working Directory > Choose Directory
knitr::opts_knit$set(root.dir = ".")
```

```{r packages}
library(UpSetR)
library(ComplexUpset)
library(ggplot2movies)
library(ggpubr)
```

```{r}
saveToPDF <- function(...) {
    d = dev.copy(pdf,...)
    dev.off(d)
}
```
# Binary list shared gene function
```{r}
fromList <- function (input) {
  # Same as original fromList()...
  elements <- unique(unlist(input))
  data <- unlist(lapply(input, function(x) {
      x <- as.vector(match(elements, x))
      }))
  data[is.na(data)] <- as.integer(0)
  data[data != 0] <- as.integer(1)
  data <- data.frame(matrix(data, ncol = length(input), byrow = F))
  data <- data[which(rowSums(data) != 0), ]
  names(data) <- names(input)
  # ... Except now it conserves your original value names!
  row.names(data) <- elements
  return(data)
  }
```

# User defined variables
```{r set_variables}
tissue <- "shared_bt_all_tissues"
control <- "Saline"
treatment <- "LPS"
control_color <- "gray29"
treatment_color <- "purple"
tool = c("star")
```

# Read in DEGs
```{r}
star_brain <- read.delim("../../../results/star/DEGs/LPS_Brain_gene_DEGs_FDRq0.05.txt", 
                    header = TRUE, sep = "\t")
star_kidney <- read.delim("../../../results/star/DEGs/LPS_Kidney_gene_DEGs_FDRq0.05.txt", 
                     header = TRUE, sep = "\t")
star_blood <- read.delim("../../../results/star/DEGs/LPS_Blood_gene_DEGs_FDRq0.05.txt", 
                     header = TRUE, sep = "\t")
```

# Subset lists
```{r}
up_star_brain <- subset(star_brain$gene_name, star_brain$logFC > 0)
down_star_brain <- subset(star_brain$gene_name, star_brain$logFC < 0)

up_star_kidney <- subset(star_kidney$gene_name, star_kidney$logFC > 0)
down_star_kidney <- subset(star_kidney$gene_name, star_kidney$logFC < 0)

up_star_blood <- subset(star_blood$gene_name, star_blood$logFC > 0)
down_star_blood <- subset(star_blood$gene_name, star_blood$logFC < 0)
```

# Upset plot between all tissues
```{r}
list_input <- list("Brain down-regulated" = down_star_brain,
                   "Brain up-regulated" = up_star_brain,
                   "Kidney down-regulated" = down_star_kidney,
                   "Kidney up-regulated" = up_star_kidney,
                   "Blood down-regulated" = down_star_blood,
                   "Blood up-regulated" = up_star_blood)
data <- fromList(list_input)

upset_gene <- upset(data, set_sizes=FALSE,
      c('Brain down-regulated','Brain up-regulated','Kidney down-regulated', 
      'Kidney up-regulated', 'Blood down-regulated', 'Blood up-regulated'),
  queries=list(
    upset_query(set='Brain down-regulated', fill='blue'),
    upset_query(set='Brain up-regulated', fill='red'),
    upset_query(set='Kidney down-regulated', fill='blue'),
    upset_query(set='Kidney up-regulated', fill='red'),
    upset_query(set='Blood down-regulated', fill='blue'),
    upset_query(set='Blood up-regulated', fill='red')
  ),
  base_annotations=list(
    'Intersection size'=(
        intersection_size(
        bar_number_threshold=1,  # show all numbers on top of bars
        width=0.5,   # reduce width of the bars
      )
      # add some space on the top of the bars
      + scale_y_continuous(expand=expansion(mult=c(0, 0.05)))
      + theme(
          # hide grid lines
          panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),
          # show axis lines
          axis.line=element_line(colour='black')
      )
    )
  ),
  stripes=upset_stripes(
    geom=geom_segment(size=12),  # make the stripes larger
    colors=c('grey95', 'white')
  ),
  # to prevent connectors from getting the colorured
  # use `fill` instead of `color`, together with `shape='circle filled'`
  matrix=intersection_matrix(
      geom=geom_point(
        shape='circle filled',
        size=3,
        stroke=0.45
      )
  ),
  sort_sets='descending',
  sort_intersections='descending'
)
upset_gene
```



# Upset plot pretty
```{r}
list_input <- list( "Blood down" = down_star_blood,
                    "Brain down" = down_star_brain,
                   "Kidney down" = down_star_kidney,
                   "Blood up" = up_star_blood, 
                   "Kidney up" = up_star_kidney,
                  "Brain up" = up_star_brain)
data <- fromList(list_input)

upset_gene <- upset(data, 
      c('Brain down','Kidney down', 'Blood down', 
        'Brain up', 'Kidney up', 'Blood up'),
      themes=upset_modify_themes(
        list('intersections_matrix'=theme(text=element_text(size=6)),
          'overall_sizes' = theme(axis.text.x = element_text(size =6)))),
  queries=list(
    upset_query(set='Brain down', fill='blue', color='blue'),
    upset_query(set='Kidney down', fill='blue', color='blue'),
    upset_query(set='Blood down', fill='blue', color='blue'),
    upset_query(set='Brain up', fill='red', color='red'),
    upset_query(set='Kidney up', fill='red', color='red'),
    upset_query(set='Blood up', fill='red', color='red'), 
    upset_query(intersect = c('Blood down','Kidney down','Brain down'), color='blue',fill='blue', only_components=c('intersections_matrix', 'Intersection size')),
    upset_query(intersect = c('Blood up', 'Kidney up','Brain up'), color='red',fill='red', only_components=c('intersections_matrix', 'Intersection size'))
  ),
  intersections =list(c('Blood down','Kidney down','Brain down'),
                      c('Brain down', 'Kidney down'), 
                      c('Brain down', 'Blood down'), 
                      c('Kidney down', 'Blood down'), 
                      'Blood down', 'Kidney down', 'Brain down', 
                      c('Blood up', 'Kidney up','Brain up'), 
                      c('Brain up', 'Kidney up'), 
                      c('Brain up', 'Blood up'), 
                      c('Kidney up', 'Blood up'), 
                      'Blood up', 'Kidney up', 'Brain up'
                      ),
  
  base_annotations=list(
    'Intersection size'=(
        intersection_size(
        size=1,
        text = list(size = 2),
        text_mapping = aes(),
        bar_number_threshold=1,  # show all numbers on top of bars
        width=0.3,   # reduce width of the bars
        mapping=aes(fill='bars_color')
      )
      + scale_fill_manual(values=c('grey'), guide='none')
      + scale_y_continuous(expand=expansion(mult=c(0, 0.1)))
      + theme(axis.text.y = element_text(size = 6),
              axis.title.y = element_text(size = 6),
          panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),
          axis.line=element_line(colour='black')))),
  matrix=intersection_matrix(
      geom=geom_point(
        shape='circle filled',
        size=2,
        stroke=0.45)) +
        theme(axis.text.y = element_text(size = 6),
              axis.title.y = element_text(size = 6)),
  set_sizes=(
    upset_set_size(geom=geom_bar(width=0.4), position="right")
    + theme(axis.line.x=element_line(colour='black'), 
            axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
            axis.title.x = element_text(size = 6),
            axis.ticks.x=element_line())),
  sort_sets=FALSE,
  sort_intersections=FALSE)
upset_gene
```
# Upset plot pretty
```{r}
list_input <- list( "Blood down" = down_star_blood,
                    "Brain down" = down_star_brain,
                   "Kidney down" = down_star_kidney,
                   "Blood up" = up_star_blood, 
                   "Kidney up" = up_star_kidney,
                  "Brain up" = up_star_brain)
data <- fromList(list_input)

upset_gene <- upset(data, 
      c('Brain down','Kidney down', 'Blood down', 
        'Brain up', 'Kidney up', 'Blood up'),
      themes=upset_modify_themes(
        list('intersections_matrix'=theme(text=element_text(size=7)),
          'overall_sizes' = theme(axis.text.x = element_text(size =7)))),
  queries=list(
    upset_query(set='Brain down', fill='blue'),
    upset_query(set='Kidney down', fill='blue'),
    upset_query(set='Blood down', fill='blue'),
    upset_query(set='Brain up', fill='red'),
    upset_query(set='Kidney up', fill='red'),
    upset_query(set='Blood up', fill='red'), 
    upset_query(intersect = c('Blood down','Kidney down','Brain down'), color='blue',fill='blue', only_components=c('intersections_matrix', 'Intersection size')),
    upset_query(intersect = c('Blood up', 'Kidney up','Brain up'), color='red',fill='red', only_components=c('intersections_matrix', 'Intersection size'))
  ),
  intersections =list(c('Blood down','Kidney down','Brain down'),
                      c('Brain down', 'Kidney down'), 
                      c('Brain down', 'Blood down'), 
                      c('Kidney down', 'Blood down'), 
                      'Blood down', 'Kidney down', 'Brain down', 
                      c('Blood up', 'Kidney up','Brain up'), 
                      c('Brain up', 'Kidney up'), 
                      c('Brain up', 'Blood up'), 
                      c('Kidney up', 'Blood up'), 
                      'Blood up', 'Kidney up', 'Brain up', 
                      c('Blood down', 'Kidney up'),
                      c('Blood down', 'Brain up'),
                      c('Blood down', 'Kidney up', 'Brain up'),
                      c('Kidney down', 'Blood up'),
                      c('Kidney down', 'Brain up'),
                  #    c('Kidney down', 'Blood up', 'Brain up') no genes
                      c('Brain down', 'Blood up'),
                      c('Brain down', 'Kidney up'),
                   #   c('Brain down', 'Blood up', 'Kidney up') no genes
                      c('Blood up', 'Kidney down', 'Brain down')
                  #    c('Kidney up', 'Blood down', 'Brain down'), no genes
                   #   c('Brain up', 'Blood down', 'Kidney down') no genes 
                      ),
  
  base_annotations=list(
    'Intersection size'=(
        intersection_size(
        size=1,
        text = list(size = 2),
        text_mapping = aes(),
        bar_number_threshold=1,  # show all numbers on top of bars
        width=0.3,   # reduce width of the bars
        mapping=aes(fill='bars_color')
      )
      + scale_fill_manual(values=c('grey'), guide='none')
      + scale_y_continuous(expand=expansion(mult=c(0, 0.1)))
      + theme(axis.text.y = element_text(size = 6),
              axis.title.y = element_text(size = 6),
          panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),
          axis.line=element_line(colour='black')))),
  matrix=intersection_matrix(
      geom=geom_point(
        shape='circle filled',
        size=2,
        stroke=0.45)) +
        theme(axis.text.y = element_text(size = 6),
              axis.title.y = element_text(size = 6)),
  set_sizes=(
    upset_set_size(geom=geom_bar(width=0.4), position="right")
    + theme(axis.line.x=element_line(colour='black'), 
            axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
            axis.title.x = element_text(size = 6),
            axis.ticks.x=element_line())),
  sort_sets=FALSE,
  sort_intersections=FALSE)
upset_gene
```

# Metascape Enrichment 
### Format the dataframes
```{r}
# read in enrichment analysis results
up_enrich_results <-
  read.delim(
    paste0(
      "../../../results/",
      tool,
      "/metascape/shared_DEGs_up.txt"
    ),
    sep = "\t",
    header = TRUE
  )

down_enrich_results <-
  read.delim(
    paste0(
      "../../../results/",
      tool,
      "/metascape/shared_DEGs_down.txt"
    ),
    sep = "\t",
    header = TRUE
  )

# select the GO term IDs we want to show in the plot
GO_ID_up <-
  c(
    "1_Summary",
    "2_Summary",
    "3_Summary",
    "4_Summary",
    "5_Summary",
    "6_Summary",
    "7_Summary",
    "8_Summary",
    "9_Summary", 
    "10_Summary",
    "11_Summary",
    "12_Summary",
    "13_Summary",
    "14_Summary",
    "15_Summary",
    "16_Summary",
    "17_Summary",
    "18_Summary",
    "19_Summary", 
    "20_Summary"
  )
GO_ID_down <- c(
    "1_Summary"
)

up_enrich_results_subset <-
  up_enrich_results[up_enrich_results$GroupID %in% GO_ID_up, ]
up_enrich_results_subset$Cluster <- c("up-regulated")
up_enrich_results_subset$Description <-
  factor(up_enrich_results_subset$Description,
         levels = up_enrich_results_subset$Description)
up_enrich_results_subset$Description <-
  fct_rev(up_enrich_results_subset$Description)

down_enrich_results_subset <-
  down_enrich_results[down_enrich_results$GroupID %in% GO_ID_down, ]
down_enrich_results_subset$Cluster <- c("down-regulated")
down_enrich_results_subset$Description <-
  factor(down_enrich_results_subset$Description,
         levels = down_enrich_results_subset$Description)
down_enrich_results_subset$Description <-
  fct_rev(down_enrich_results_subset$Description)

# get the number of genes in each summary
up_gene_count <-
  strsplit(as.character(up_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
up_gene_count_df <-
  data.frame(matrix(
    unlist(up_gene_count),
    nrow = length(up_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
up_enrich_results_subset$InTerm <- as.numeric(up_gene_count_df$X1)
up_enrich_results_subset$InList <- as.numeric(up_gene_count_df$X2)

down_gene_count <-
  strsplit(as.character(down_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
down_gene_count_df <-
  data.frame(matrix(
    unlist(down_gene_count),
    nrow = length(down_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
down_enrich_results_subset$InTerm <- as.numeric(down_gene_count_df$X1)
down_enrich_results_subset$InList <- as.numeric(down_gene_count_df$X2)

# combine together
up_and_down_enrich_results_subset <- rbind(up_enrich_results_subset, down_enrich_results_subset)

#up_and_down_enrich_results_subset$Description <-
 #     gsub(
 #       "regulation of plasma membrane bounded cell projection organization",
#        "reg of plasma membrane/cell projection organization",
 #       up_and_down_enrich_results_subset$Description
  #    )
up_and_down_enrich_results_subset$Description <-
  factor(up_and_down_enrich_results_subset$Description, levels = unique(up_and_down_enrich_results_subset$Description)) 

up_and_down_enrich_results_subset$Description <- fct_rev(up_and_down_enrich_results_subset$Description)
```

remove files 
```{r}
remove(
  down_enrich_results,
  down_gene_count,
  down_gene_count_df,
  up_enrich_results,
  up_gene_count,
  up_gene_count_df
)
```

### Enrichment plot showing log10 P-value
#### down and up seperately
```{r}
down_enrich_plot <-
  ggplot(data = down_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster) +
  geom_bar(stat = "identity", aes(fill = LogP), width = .7, position = position_dodge(width = .2)) +
 ggtitle("shared down DEGs enrichment summaries") +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = "Gene count", y = "") +
  #  xlim(0,60) +
  guides(fill = guide_legend(
    size = 1,
    title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)")
  )) +
  scale_fill_gradientn(
    colours = c("darkmagenta", "mediumorchid3", "grey"),
    guide = "legend",
    limits = c(-45,0),
  ) +
    theme(strip.text = element_text(size = 6), 
          axis.text.y = element_text(size = 7),
          axis.text.x = element_text(size = 6), 
          axis.title.x = element_text(size = 6), 
          legend.text = element_text(size = 6),
          legend.title = element_text(size = 6),
          legend.margin=margin(0,0.5,0,0),
          legend.box.margin=margin(-10,-2,-10,-7.5), 
          plot.margin = margin(0.1, 0.2, 0, 0.2, "cm"), 
          plot.title = element_text(size = 7, hjust = -3.25, vjust=0, margin = margin(0,0,0,0)))
down_enrich_plot

up_enrich_plot <-
  ggplot(data = up_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster) +
  geom_bar(stat = "identity", aes(fill = LogP), width = .7, position = position_dodge(width = .2)) +
  ggtitle("shared up DEGs enrichment summaries") +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = "Gene count", y = "") +
  guides(fill = guide_legend(title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)"))) +
  scale_fill_gradientn(
    colours = c("darkmagenta", "mediumorchid3", "grey"),
    guide = "legend",
    limits = c(-45,-1)
  ) +
    theme(strip.text = element_text(size = 6), 
          axis.text.y = element_text(size = 7),
          axis.text.x = element_text(size = 6), 
          axis.title.x = element_text(size = 6), 
          legend.text = element_text(size = 6),
          legend.title = element_text(size = 6),
          legend.margin=margin(0,0.5,0,0),
          legend.box.margin=margin(-10,-2,-10,-7.5), 
          plot.margin = margin(0.1, 0.2, 0, 0.2, "cm"), 
          plot.title = element_text(size = 7, hjust = -3.25, vjust=0, margin = margin(0,0,0,0)))
up_enrich_plot
```

#### down and up together
```{r}
up_and_down_enrich_plot <-
  ggplot(data = up_and_down_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster, scales = "free") +
  geom_bar(stat = "identity", aes(fill = LogP), width = .7, position = position_dodge(width = .2)) +
  theme_bw() +
  ggtitle(paste0("Shared DEGs enrichment summaries")) +
  labs(x = "Gene count", y = NULL) +
  guides(fill = guide_legend(title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)"))) +
  scale_fill_gradientn(
    colours = c("darkmagenta", "mediumorchid3", "grey"),
    guide = "legend",
    limits = c(-45,-1)
  ) +
    theme(strip.text = element_text(size = 6), 
          axis.text.y = element_text(size = 7),
          axis.text.x = element_text(size = 6), 
          axis.title.x = element_text(size = 6), 
          legend.text = element_text(size = 6),
          legend.title = element_text(size = 6),
          legend.margin=margin(0,0.5,0,0),
          legend.box.margin=margin(-10,-2,-10,-7.5), 
          plot.margin = margin(0.1, 0.2, 0, 0.2, "cm"), 
          plot.title = element_text(size = 7, hjust = -3.25, vjust=0, margin = margin(0,0,1,0)))
up_and_down_enrich_plot

addSmallLegend <- function(myPlot, pointSize = 3, textSize = 6, spaceLegend = .5) {
    myPlot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}

# Apply on original plot
up_and_down_enrich_plot <- addSmallLegend(up_and_down_enrich_plot)
```

# Heatmap of fold change for each tissue
read in treatment vs control DEG table for each tissue
read in lcpm data for each tissue 
%in% to limit to the list of genes that are in the metascape analysis 
melt
order by fold change
```{r}
# brain
tissue = "Brain"
treatment_vs_control <- read.table(
  paste0(
    "../../../results/",
    tool,
    "/DEGs/",
    treatment,
    "_",
    tissue,
    "_gene_DEGs_FDRq1.00.txt",
    sep = ""
  ),
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)

up_regulated <- subset(
  treatment_vs_control,
  treatment_vs_control$adj.P.Val <= 0.05 &
    treatment_vs_control$logFC > 0
)
down_regulated <- subset(
  treatment_vs_control,
  treatment_vs_control$adj.P.Val <= 0.05 &
    treatment_vs_control$logFC <= 0
)

# sort by log2FC and adjusted p-val
up_regulated_sort <-
  up_regulated[order(-up_regulated$logFC, up_regulated$adj.P.Val), ]
down_regulated_sort <-
  down_regulated[order(down_regulated$logFC, down_regulated$adj.P.Val), ]
up_regulated_sort$tissue <- c("Brain")
down_regulated_sort$tissue <- c("Brain")


# lock in gene order which is sorted by adjusted p-value
up_regulated_sort$gene<-
  factor(up_regulated_sort$gene_name,
         levels = unique(up_regulated_sort$gene_name))
up_regulated_sort$gene <-
  fct_rev(up_regulated_sort$gene)
down_regulated_sort$gene <-
  factor(down_regulated_sort$gene_name,
         levels = unique(down_regulated_sort$gene_name))
down_regulated_sort$gene <-
  fct_rev(down_regulated_sort$gene)

brain_FC_up <- data.frame(up_regulated_sort$gene, up_regulated_sort$tissue, up_regulated_sort$logFC)
brain_FC_down <- data.frame(down_regulated_sort$gene, down_regulated_sort$tissue, down_regulated_sort$logFC)

#remove files
remove(down_regulated_sort,
  down_regulated,
  up_regulated,
  up_regulated_sort, 
  treatment_vs_control
)

# kidney
tissue = "Kidney"
treatment_vs_control <- read.table(
  paste0(
    "../../../results/",
    tool,
    "/DEGs/",
    treatment,
    "_",
    tissue,
    "_gene_DEGs_FDRq1.00.txt",
    sep = ""
  ),
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)

up_regulated <- subset(
  treatment_vs_control,
  treatment_vs_control$adj.P.Val <= 0.05 &
    treatment_vs_control$logFC > 0
)
down_regulated <- subset(
  treatment_vs_control,
  treatment_vs_control$adj.P.Val <= 0.05 &
    treatment_vs_control$logFC <= 0
)

# sort by log2FC and adjusted p-val
up_regulated_sort <-
  up_regulated[order(-up_regulated$logFC, up_regulated$adj.P.Val), ]
down_regulated_sort <-
  down_regulated[order(down_regulated$logFC, down_regulated$adj.P.Val), ]
up_regulated_sort$tissue <- c("Kidney")
down_regulated_sort$tissue <- c("Kidney")


# lock in gene order which is sorted by adjusted p-value
up_regulated_sort$gene<-
  factor(up_regulated_sort$gene_name,
         levels = unique(up_regulated_sort$gene_name))
up_regulated_sort$gene <-
  fct_rev(up_regulated_sort$gene)
down_regulated_sort$gene <-
  factor(down_regulated_sort$gene_name,
         levels = unique(down_regulated_sort$gene_name))
down_regulated_sort$gene <-
  fct_rev(down_regulated_sort$gene)

kidney_FC_up <- data.frame(up_regulated_sort$gene, up_regulated_sort$tissue, up_regulated_sort$logFC)
kidney_FC_down <- data.frame(down_regulated_sort$gene, down_regulated_sort$tissue, down_regulated_sort$logFC)

#remove files
remove(down_regulated_sort,
  down_regulated,
  up_regulated,
  up_regulated_sort,
  treatment_vs_control
)

# Blood
tissue = "Blood"
treatment_vs_control <- read.table(
  paste0(
    "../../../results/",
    tool,
    "/DEGs/",
    treatment,
    "_",
    tissue,
    "_gene_DEGs_FDRq1.00.txt",
    sep = ""
  ),
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)

up_regulated <- subset(
  treatment_vs_control,
  treatment_vs_control$adj.P.Val <= 0.05 &
    treatment_vs_control$logFC > 0
)
down_regulated <- subset(
  treatment_vs_control,
  treatment_vs_control$adj.P.Val <= 0.05 &
    treatment_vs_control$logFC <= 0
)

# sort by log2FC and adjusted p-val
up_regulated_sort <-
  up_regulated[order(-up_regulated$logFC, up_regulated$adj.P.Val), ]
down_regulated_sort <-
  down_regulated[order(down_regulated$logFC, down_regulated$adj.P.Val), ]
up_regulated_sort$tissue <- c("Blood")
down_regulated_sort$tissue <- c("Blood")


# lock in gene order which is sorted by adjusted p-value
up_regulated_sort$gene<-
  factor(up_regulated_sort$gene_name,
         levels = unique(up_regulated_sort$gene_name))
up_regulated_sort$gene <-
  fct_rev(up_regulated_sort$gene)
down_regulated_sort$gene <-
  factor(down_regulated_sort$gene_name,
         levels = unique(down_regulated_sort$gene_name))
down_regulated_sort$gene <-
  fct_rev(down_regulated_sort$gene)

blood_FC_up <- data.frame(up_regulated_sort$gene, up_regulated_sort$tissue, up_regulated_sort$logFC)
blood_FC_down <- data.frame(down_regulated_sort$gene, down_regulated_sort$tissue, down_regulated_sort$logFC)

#remove files
remove(down_regulated_sort,
  down_regulated,
  up_regulated,
  up_regulated_sort, 
  treatment_vs_control
)

# rename columns and bind together
names(brain_FC_down)[1:3] <- c('gene', 'tissue', 'logFC')
names(brain_FC_up)[1:3] <- c('gene', 'tissue', 'logFC')
names(kidney_FC_down)[1:3] <- c('gene', 'tissue', 'logFC')
names(kidney_FC_up)[1:3] <- c('gene', 'tissue', 'logFC')
names(blood_FC_down)[1:3] <- c('gene', 'tissue', 'logFC')
names(blood_FC_up)[1:3] <- c('gene', 'tissue', 'logFC')

up_FC <- rbind(brain_FC_up, kidney_FC_up, blood_FC_up)
down_FC <- rbind(brain_FC_down, kidney_FC_down, blood_FC_down)
```

### Genes in enrichment plot all
## up 

Cytokine Signaling in Immune system	[1]
TNF signaling pathway	[4]
 [7] positive regulation of cytokine production  
[10] PID IL12 2PATHWAY  
[16] regulation of I-kappaB kinase/NF-kappaB signaling 
[17] inflammatory response 
```{r}
up_enrich_df <-
  up_enrich_results_subset[, c("GroupID", "Term", "Description", "Symbols")]
up_enrich_df_melt <- reshape2::melt(up_enrich_df)
up_gene_char <- str_split(up_enrich_df_melt$Symbols, ",")

up_gene_pathway <-
  data.frame(gene = unlist(up_gene_char),
             value = rep(up_enrich_df$Description, lengths(up_gene_char)))
up_pathways_levels <- levels(fct_rev(up_gene_pathway$value))

up_input <- list(
  "Cytokine Signaling in Immune system" = up_gene_char[[1]],
  "TNF signaling pathway" = up_gene_char[[5]],
  "positive regulation of cytokine production" = up_gene_char[[7]],
  "PID IL12 2PATHWAY" = up_gene_char[[10]],
  "regulation of I-kappaB kinase/NF-kappaB signaling" = up_gene_char[[16]]
)
up_data <- fromList(up_input)
up_data$gene <- row.names(up_data)
up_data_melt <- reshape2::melt(up_data)


# get the fold change value for those genes
up_df <- merge(up_data_melt, up_FC,
                       by = "gene")
# order by fold change 
up_df <- up_df[order(-up_df$logFC),]
up_df$tissue <- factor(up_df$tissue, levels = c("Blood", "Kidney", "Brain"))

# subset by GO termes
GO_1 <-
  subset(up_df,
         up_df$variable == "Cytokine Signaling in Immune system"
         & up_df$value == 1)
GO_1_top <- head(GO_1, 35)

GO_2 <-
  subset(up_df,
         up_df$variable == "TNF signaling pathway"
         & up_df$value == 1)
GO_2_top <- head(GO_2, 60)

GO_3 <-
  subset(up_df,
         up_df$variable == "positive regulation of cytokine production"
         & up_df$value == 1)
GO_3_top <- head(GO_3, 45)

GO_4 <-
  subset(up_df,
         up_df$variable == "PID IL12 2PATHWAY"
         & up_df$value == 1)
GO_4_top <- head(GO_4, 35)

GO_5 <-
  subset(up_df,
         up_df$variable == "regulation of I-kappaB kinase/NF-kappaB signaling"
         & up_df$value == 1)
GO_5_top <- head(GO_5, 45)
```
### up 1 heatmap 
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_1_DEG <- subset(up_df, up_df$gene %in% GO_1_top$gene)
GO_1_DEG$gene <- factor(GO_1_DEG$gene,
                          levels = unique(GO_1_top$gene))
GO_1_DEG$gene <- fct_rev(GO_1_DEG$gene)
# heatmap 
up_GO_1_lcpm_heat <- ggplot(data = GO_1_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "white",
    mid = "red",
    high = "maroon",
    midpoint = 4.5,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 3, 6, 9),
    limits = c(0,9),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0.2, 0.5, 0.2, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("Cytokine Signaling in 
Immune system") 
up_GO_1_lcpm_heat
```

### up 2 heatmap 
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_2_DEG <- subset(up_df, up_df$gene %in% GO_2_top$gene)
GO_2_DEG$gene <- factor(GO_2_DEG$gene,
                          levels = unique(GO_2_top$gene))
GO_2_DEG$gene <- fct_rev(GO_2_DEG$gene)
# heatmap 
up_GO_2_lcpm_heat <- ggplot(data = GO_2_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "white",
    mid = "red",
    high = "maroon",
    midpoint = 4.5,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 3, 6, 9),
    limits = c(0,9),
    name = expression(log[2](FC))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0.2, 0.5, 0.2, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("TNF signaling 
pathway") 
up_GO_2_lcpm_heat
```
### up 3 heatmap 
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_3_DEG <- subset(up_df, up_df$gene %in% GO_3_top$gene)
GO_3_DEG$gene <- factor(GO_3_DEG$gene,
                          levels = unique(GO_3_top$gene))
GO_3_DEG$gene <- fct_rev(GO_3_DEG$gene)
# heatmap 
up_GO_3_lcpm_heat <- ggplot(data = GO_3_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "white",
    mid = "red",
    high = "maroon",
    midpoint = 4.5,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 3, 6, 9),
    limits = c(0,9),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0.2, 0.5, 0.2, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("positive regulation 
of cytokine 
production") 
up_GO_3_lcpm_heat
```

### up 4 heatmap 
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_4_DEG <- subset(up_df, up_df$gene %in% GO_4_top$gene)
GO_4_DEG$gene <- factor(GO_4_DEG$gene,
                          levels = unique(GO_4_top$gene))
GO_4_DEG$gene <- fct_rev(GO_4_DEG$gene)
# heatmap 
up_GO_4_lcpm_heat <- ggplot(data = GO_4_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "white",
    mid = "red",
    high = "maroon",
    midpoint = 4.5,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 3, 6, 9),
    limits = c(0,9),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0.2, 0.5, 0.2, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("PID IL12 
2PATHWAY") 
up_GO_4_lcpm_heat
```

### up 5 heatmap 
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_5_DEG <- subset(up_df, up_df$gene %in% GO_5_top$gene)
GO_5_DEG$gene <- factor(GO_5_DEG$gene,
                          levels = unique(GO_5_top$gene))
GO_5_DEG$gene <- fct_rev(GO_5_DEG$gene)
# heatmap 
up_GO_5_lcpm_heat <- ggplot(data = GO_5_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "white",
    mid = "red",
    high = "maroon",
    midpoint = 4.5,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 3, 6, 9),
    limits = c(0,9),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0.2, 0.5, 0.2, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("regulation of 
I-kappaB
kinase/NF-kappaB 
signaling") 
up_GO_5_lcpm_heat
```
## up heatmap legend
```{r}
legend_data_up <- ggplot(data = GO_2_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "white",
    mid = "red",
    high = "maroon",
    midpoint = 4.5,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 3, 6, 9),
    limits = c(0,9),
    name = expression(log[2](FC))
  ) +
  theme(
    legend.position = "top", 
    legend.text = element_text(size =6), 
    legend.title = element_text (size = 6), 
    legend.margin=margin(0,0,1,0),
  #  legend.direction = "horizontal", 
    legend.key.size = unit(0.5, "cm"))

legend_data_up <- addSmallLegend(legend_data_up)

leg_up <- get_legend(legend_data_up)
# Convert to a ggplot and print
legend_tpm_bar_up <- as_ggplot(leg_up)
legend_tpm_bar_up
```
## down 

```{r}
down_enrich_df <-
  down_enrich_results_subset[, c("GroupID", "Term", "Description", "Symbols")]
down_enrich_df_melt <- reshape2::melt(down_enrich_df)
down_gene_char <- str_split(down_enrich_df_melt$Symbols, ",")

down_gene_pathway <-
  data.frame(gene = unlist(down_gene_char),
             value = rep(down_enrich_df$Description, lengths(down_gene_char)))
down_pathways_levels <- levels(fct_rev(down_gene_pathway$value))

down_input <- list(
  "RHO GTPase cycle" = down_gene_char[[1]], 
  "RHO GTPase cycle" = down_gene_char[[1]]
)
down_data <- fromList(down_input)
down_data$gene <- row.names(down_data)
down_data_melt <- reshape2::melt(down_data)


# get the fold change value for those genes
down_df <- merge(down_data_melt, down_FC,
                       by = "gene")
# order by fold change 
down_df <- down_df[order(-down_df$logFC),]
down_df$tissue <- factor(down_df$tissue, levels = c("Blood", "Kidney", "Brain"))

# subset by GO termes
GO_1 <-
  subset(down_df,
         down_df$variable == "RHO GTPase cycle"
         & down_df$value == 1)
GO_1_top <- head(GO_1, 45)
```

### down 1 heatmap 
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_1_DEG <- subset(down_df, down_df$gene %in% GO_1_top$gene)
GO_1_DEG$gene <- factor(GO_1_DEG$gene,
                          levels = unique(GO_1_top$gene))
GO_1_DEG$gene <- fct_rev(GO_1_DEG$gene)
# heatmap 
down_GO_1_lcpm_heat <- ggplot(data = GO_1_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "navyblue",
    mid = "blue",
    high = "white",
    midpoint = -2,
    space = "rgb",
    guide = "colourbar",
    breaks = c(-4, -3, -2, -1, 0),
    limits = c(-4,0),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0.2, 0.5, 0.2, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("RHO GTPase 
cycle") 
down_GO_1_lcpm_heat
```

## down heatmap legend
```{r}
# heatmap 
legend_data_down <- ggplot(data = GO_2_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "navyblue",
    mid = "blue",
    high = "white",
    midpoint = -2,
    space = "rgb",
    guide = "colourbar",
    breaks = c(-4,-3,-2,-1,0),
    limits = c(-4,0),
    name = expression(log[2](FC))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme(
    legend.position = "top", 
    legend.text = element_text(size =6), 
    legend.title = element_text (size = 6), 
    legend.margin=margin(0,0,1,0),
   # legend.direction = "vertical", 
    legend.key.size = unit(0.5, "cm"))

legend_data_down <- addSmallLegend(legend_data_down)

leg_down <- get_legend(legend_data_down)
# Convert to a ggplot and print
legend_tpm_bar_down <- as_ggplot(leg_down)
legend_tpm_bar_down
```


# Combine volcano and enrichment plot 
```{r}
row1 <- ggarrange(
  upset_gene,
  labels = c("a)"),
  font.label = list(size = 8)
  )
row1

row2.1 <- ggarrange(
  up_and_down_enrich_plot,
  labels = c("b)"),
  font.label = list(size = 8))

row2.2 <-
  ggarrange(
    NULL,
    legend_tpm_bar_down, 
    down_GO_1_lcpm_heat,
    NULL, 
    labels = c("","", "c)"),
    nrow = 4, 
    heights = c(0.5, 0.5, 1, 0.5),
    font.label = list(size = 8, color = "black")
  )

row2 <- ggarrange(
  row2.1,
  row2.2,
  ncol = 2,
  widths = c(2,0.65)
)

row3 <-
  ggarrange(
    NULL, 
    legend_tpm_bar_up,
    NULL,
    NULL,
    NULL,
    ncol = 5,
    labels = c("up-regulated", ""),
    widths = c(1,0.15, 1, 1,1),
    font.label = list(size = 7, color = "red")
  )

row4 <- ggarrange(
  up_GO_1_lcpm_heat, 
  up_GO_2_lcpm_heat, 
  up_GO_4_lcpm_heat, 
  up_GO_5_lcpm_heat, 
  nrow = 1, 
  ncol = 4,
  labels = c("d)", "e)", "f)", "g)"),
  font.label = list(size = 8)
)


together <- ggarrange(row1, row2, row3, row4, nrow = 4, heights = c(2.35, 2, 0.25,2))
together
path <- paste0("../../../results/manuscript_figures/Figure_",
               "shared_DEGs",
               "_gene_level_only")
saveToPDF(paste0(path, ".pdf"), width = 7.08, height = 8.66)
```
# session info
```{r}
```
