---
title: "Supplemental figure kidney brain heatmaps"
author: "Kimberly Olney"
date: "6/14/2022"
output: html_document
---

# Setup
```{r setup, include=FALSE}
# Also do Session > Set Working Directory > Choose Directory
knitr::opts_knit$set(root.dir = ".")
```

```{r packages}
library(UpSetR)
library(ComplexUpset)
library(ggplot2movies)
library(ggpubr)
library(forcats)
library(stringr)
```

```{r}
saveToPDF <- function(...) {
    d = dev.copy(pdf,...)
    dev.off(d)
}
```
# Binary list shared gene function
```{r}
fromList <- function (input) {
  # Same as original fromList()...
  elements <- unique(unlist(input))
  data <- unlist(lapply(input, function(x) {
      x <- as.vector(match(elements, x))
      }))
  data[is.na(data)] <- as.integer(0)
  data[data != 0] <- as.integer(1)
  data <- data.frame(matrix(data, ncol = length(input), byrow = F))
  data <- data[which(rowSums(data) != 0), ]
  names(data) <- names(input)
  # ... Except now it conserves your original value names!
  row.names(data) <- elements
  return(data)
  }
```

# User defined variables
```{r set_variables}
tissue <- "shared_bt_brain_and_kidney"
control <- "Saline"
treatment <- "LPS"
control_color <- "gray29"
treatment_color <- "purple"
tool = c("star")
```

# Read in DEGs
```{r}
star_brain <- read.delim("../../../results/star/DEGs/LPS_Brain_gene_DEGs_FDRq0.05.txt", 
                    header = TRUE, sep = "\t")
star_kidney <- read.delim("../../../results/star/DEGs/LPS_Kidney_gene_DEGs_FDRq0.05.txt", 
                     header = TRUE, sep = "\t")
```

# Subset lists
```{r}
up_star_brain <- subset(star_brain$gene_name, star_brain$logFC > 0)
down_star_brain <- subset(star_brain$gene_name, star_brain$logFC < 0)

up_star_kidney <- subset(star_kidney$gene_name, star_kidney$logFC > 0)
down_star_kidney <- subset(star_kidney$gene_name, star_kidney$logFC < 0)
```

# Upset plot between all tissues
```{r}
list_input <- list("Brain down-regulated" = down_star_brain,
                   "Brain up-regulated" = up_star_brain,
                   "Kidney down-regulated" = down_star_kidney,
                   "Kidney up-regulated" = up_star_kidney)
data <- fromList(list_input)

upset_gene <- upset(data, set_sizes=FALSE,
      c('Brain down-regulated','Brain up-regulated','Kidney down-regulated', 
      'Kidney up-regulated'),
  queries=list(
    upset_query(set='Brain down-regulated', fill='blue'),
    upset_query(set='Brain up-regulated', fill='red'),
    upset_query(set='Kidney down-regulated', fill='blue'),
    upset_query(set='Kidney up-regulated', fill='red')
  ),
  base_annotations=list(
    'Intersection size'=(
        intersection_size(
        bar_number_threshold=1,  # show all numbers on top of bars
        width=0.5,   # reduce width of the bars
      )
      # add some space on the top of the bars
      + scale_y_continuous(expand=expansion(mult=c(0, 0.05)))
      + theme(
          # hide grid lines
          panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),
          # show axis lines
          axis.line=element_line(colour='black')
      )
    )
  ),
  stripes=upset_stripes(
    geom=geom_segment(size=12),  # make the stripes larger
    colors=c('grey95', 'white')
  ),
  # to prevent connectors from getting the colorured
  # use `fill` instead of `color`, together with `shape='circle filled'`
  matrix=intersection_matrix(
      geom=geom_point(
        shape='circle filled',
        size=3,
        stroke=0.45
      )
  ),
  sort_sets='descending',
  sort_intersections='descending'
)
upset_gene
```

# Upset plot pretty
```{r}
list_input <- list("Brain down" = down_star_brain,
                   "Kidney down" = down_star_kidney,
                   "Kidney up" = up_star_kidney,
                  "Brain up" = up_star_brain)
data <- fromList(list_input)

upset_gene <- upset(data, 
      c('Brain down','Kidney down',
        'Brain up', 'Kidney up'),
      themes=upset_modify_themes(
        list('intersections_matrix'=theme(text=element_text(size=6)),
          'overall_sizes' = theme(axis.text.x = element_text(size =6)))),
  queries=list(
    upset_query(set='Brain down', fill='blue'),
    upset_query(set='Kidney down', fill='blue'),
    upset_query(set='Brain up', fill='red'),
    upset_query(set='Kidney up', fill='red'),
    upset_query(intersect = c('Kidney down','Brain down'), color='blue',fill='blue', only_components=c('intersections_matrix', 'Intersection size')),
    upset_query(intersect = c('Kidney up','Brain up'), color='red',fill='red', only_components=c('intersections_matrix', 'Intersection size'))
  ),
  intersections =list(c('Brain down', 'Kidney down'), 
                      'Kidney down', 'Brain down', 
                      c('Brain up', 'Kidney up'), 
                      'Kidney up', 'Brain up',
                      c('Brain down', 'Kidney up'), 
                      c('Brain up', 'Kidney down')
                      ),
  
  base_annotations=list(
    'Intersection size'=(
        intersection_size(
        size=1,
        text = list(size = 2),
        text_mapping = aes(),
        bar_number_threshold=1,  # show all numbers on top of bars
        width=0.3,   # reduce width of the bars
        mapping=aes(fill='bars_color')
      )
      + scale_fill_manual(values=c('grey'), guide='none')
      + scale_y_continuous(expand=expansion(mult=c(0, 0.1)))
      + theme(axis.text.y = element_text(size = 6),
              axis.title.y = element_text(size = 6),
          panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),
          axis.line=element_line(colour='black')))),
  matrix=intersection_matrix(
      geom=geom_point(
        shape='circle filled',
        size=2,
        stroke=0.45)) +
        theme(axis.text.y = element_text(size = 6),
              axis.title.y = element_text(size = 6)),
  set_sizes=(
    upset_set_size(geom=geom_bar(width=0.4), position="right")
    + theme(axis.line.x=element_line(colour='black'), 
            axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
            axis.title.x = element_text(size = 6),
            axis.ticks.x=element_line())),
  sort_sets=FALSE,
  sort_intersections=FALSE)
upset_gene
```

# Metascape Enrichment 
### Format the dataframes
```{r}
# read in enrichment analysis results
up_enrich_results <-
  read.delim(
    paste0(
      "../../../results/",
      tool,
      "/metascape/shared_DEGs_brain_kidney_up.txt"
    ),
    sep = "\t",
    header = TRUE
  )

down_enrich_results <-
  read.delim(
    paste0(
      "../../../results/",
      tool,
      "/metascape/shared_DEGs_brain_kidney_down.txt"
    ),
    sep = "\t",
    header = TRUE
  )

# select the GO term IDs we want to show in the plot
GO_ID_up <-
  c(
    "1_Summary",
    "2_Summary",
    "3_Summary",
    "4_Summary",
    "5_Summary",
    "6_Summary",
    "7_Summary",
    "8_Summary",
    "9_Summary", 
    "10_Summary",
    "11_Summary",
    "12_Summary",
    "13_Summary",
    "14_Summary",
    "15_Summary",
    "16_Summary",
    "17_Summary",
    "18_Summary",
    "19_Summary", 
    "20_Summary"
  )
GO_ID_down <- c(
    "1_Summary",
    "2_Summary",
    "3_Summary",
    "4_Summary",
    "5_Summary",
    "6_Summary",
    "7_Summary",
    "8_Summary",
    "9_Summary", 
    "10_Summary",
    "11_Summary"
)

up_enrich_results_subset <-
  up_enrich_results[up_enrich_results$GroupID %in% GO_ID_up, ]
up_enrich_results_subset$Cluster <- c("up-regulated")
up_enrich_results_subset$Description <-
  factor(up_enrich_results_subset$Description,
         levels = up_enrich_results_subset$Description)
up_enrich_results_subset$Description <-
  fct_rev(up_enrich_results_subset$Description)

down_enrich_results_subset <-
  down_enrich_results[down_enrich_results$GroupID %in% GO_ID_down, ]
down_enrich_results_subset$Cluster <- c("down-regulated")
down_enrich_results_subset$Description <-
  factor(down_enrich_results_subset$Description,
         levels = down_enrich_results_subset$Description)
down_enrich_results_subset$Description <-
  fct_rev(down_enrich_results_subset$Description)

# get the number of genes in each summary
up_gene_count <-
  strsplit(as.character(up_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
up_gene_count_df <-
  data.frame(matrix(
    unlist(up_gene_count),
    nrow = length(up_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
up_enrich_results_subset$InTerm <- as.numeric(up_gene_count_df$X1)
up_enrich_results_subset$InList <- as.numeric(up_gene_count_df$X2)

down_gene_count <-
  strsplit(as.character(down_enrich_results_subset$InTerm_InList),
           "/",
           fixed = T)
down_gene_count_df <-
  data.frame(matrix(
    unlist(down_gene_count),
    nrow = length(down_gene_count),
    byrow = TRUE
  ), stringsAsFactors = FALSE)
down_enrich_results_subset$InTerm <- as.numeric(down_gene_count_df$X1)
down_enrich_results_subset$InList <- as.numeric(down_gene_count_df$X2)

# combine together
up_and_down_enrich_results_subset <- rbind(up_enrich_results_subset, down_enrich_results_subset)

up_and_down_enrich_results_subset$Description <- gsub(
  "TNF-related weak inducer of apoptosis (TWEAK) signaling pathway",
  "TNF weak inducer of apoptosis signaling pathway", fixed=TRUE,
  up_and_down_enrich_results_subset$Description)


up_and_down_enrich_results_subset$Description <-
  factor(up_and_down_enrich_results_subset$Description, levels = unique(up_and_down_enrich_results_subset$Description)) 

up_and_down_enrich_results_subset$Description <- fct_rev(up_and_down_enrich_results_subset$Description)
```

remove files 
```{r}
remove(
  down_enrich_results,
  down_gene_count,
  down_gene_count_df,
  up_enrich_results,
  up_gene_count,
  up_gene_count_df
)
```

### Enrichment plot showing log10 P-value
#### down and up seperately
```{r}
down_enrich_plot <-
  ggplot(data = down_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster) +
  geom_bar(stat = "identity", aes(fill = LogP), width = .7, position = position_dodge(width = .2)) +
 ggtitle("shared down DEGs enrichment summaries") +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = "Gene count", y = "") +
  #  xlim(0,60) +
  guides(fill = guide_legend(
    size = 1,
    title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)")
  )) +
  scale_fill_gradientn(
    colours = c("darkmagenta", "mediumorchid3", "grey"),
    guide = "legend",
    limits = c(-45,0),
  ) +
    theme(strip.text = element_text(size = 6), 
          axis.text.y = element_text(size = 7),
          axis.text.x = element_text(size = 6), 
          axis.title.x = element_text(size = 6), 
          legend.text = element_text(size = 6),
          legend.title = element_text(size = 6),
          legend.margin=margin(0,0.5,0,0),
          legend.box.margin=margin(-10,-2,-10,-7.5), 
          plot.margin = margin(0.1, 0.2, 0, 0.2, "cm"), 
          plot.title = element_text(size = 7, hjust = 1, vjust=0, margin = margin(0,0,0,0)))
down_enrich_plot

up_enrich_plot <-
  ggplot(data = up_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster) +
  geom_bar(stat = "identity", aes(fill = LogP), width = .7, position = position_dodge(width = .2)) +
  ggtitle("shared up DEGs enrichment summaries") +
  scale_y_discrete(expand = c(0, 0)) +
  labs(x = "Gene count", y = "") +
  guides(fill = guide_legend(title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)"))) +
  scale_fill_gradientn(
    colours = c("darkmagenta", "mediumorchid3", "grey"),
    guide = "legend",
    limits = c(-45,-1)
  ) +
    theme(strip.text = element_text(size = 6), 
          axis.text.y = element_text(size = 7),
          axis.text.x = element_text(size = 6), 
          axis.title.x = element_text(size = 6), 
          legend.text = element_text(size = 6),
          legend.title = element_text(size = 6),
          legend.margin=margin(0,0.5,0,0),
          legend.box.margin=margin(-10,-2,-10,-7.5), 
          plot.margin = margin(0.1, 0.2, 0, 0.2, "cm"), 
          plot.title = element_text(size = 7, hjust = -3.25, vjust=0, margin = margin(0,0,0,0)))
up_enrich_plot
```

#### down and up together
```{r}
up_and_down_enrich_plot <-
  ggplot(data = up_and_down_enrich_results_subset, aes(x = InTerm, y = Description)) +
  ggplot2::facet_grid(~ Cluster, scales = "free") +
  geom_bar(stat = "identity", aes(fill = LogP), width = .7, position = position_dodge(width = .2)) +
  theme_bw() +
  ggtitle(paste0("Shared Brain and Kidney DEGs enrichment summaries")) +
  labs(x = "Gene count", y = NULL) +
  guides(fill = guide_legend(title = expression(log[10] ~ "(" ~ italic("p") ~ "-value)"))) +
  scale_fill_gradientn(
    colours = c("darkmagenta", "mediumorchid3", "grey"),
    guide = "legend",
    limits = c(-45,-1)
  ) +
    theme(strip.text = element_text(size = 6), 
          axis.text.y = element_text(size = 7),
          axis.text.x = element_text(size = 6), 
          axis.title.x = element_text(size = 6), 
          legend.text = element_text(size = 6),
          legend.title = element_text(size = 6),
          legend.margin=margin(0,0.5,0,0),
          legend.box.margin=margin(-10,-2,-10,-7.5), 
          plot.margin = margin(0.1, 0.2, 0, 0.2, "cm"), 
          plot.title = element_text(size = 7, hjust = -1.65, vjust=0, margin = margin(0,0,1,0)))
up_and_down_enrich_plot

addSmallLegend <- function(myPlot, pointSize = 3, textSize = 6, spaceLegend = .5) {
    myPlot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}

# Apply on original plot
up_and_down_enrich_plot <- addSmallLegend(up_and_down_enrich_plot)
```

# Heatmap of fold change for each tissue
read in treatment vs control DEG table for each tissue
read in lcpm data for each tissue 
%in% to limit to the list of genes that are in the metascape analysis 
melt
order by fold change
```{r}
# brain
tissue = "Brain"
treatment_vs_control <- read.table(
  paste0(
    "../../../results/",
    tool,
    "/DEGs/",
    treatment,
    "_",
    tissue,
    "_gene_DEGs_FDRq1.00.txt",
    sep = ""
  ),
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)

up_regulated <- subset(
  treatment_vs_control,
  treatment_vs_control$adj.P.Val <= 0.05 &
    treatment_vs_control$logFC > 0
)
down_regulated <- subset(
  treatment_vs_control,
  treatment_vs_control$adj.P.Val <= 0.05 &
    treatment_vs_control$logFC <= 0
)

# sort by log2FC and adjusted p-val
up_regulated_sort <-
  up_regulated[order(-up_regulated$logFC, up_regulated$adj.P.Val), ]
down_regulated_sort <-
  down_regulated[order(down_regulated$logFC, down_regulated$adj.P.Val), ]
up_regulated_sort$tissue <- c("Brain")
down_regulated_sort$tissue <- c("Brain")


# lock in gene order which is sorted by adjusted p-value
up_regulated_sort$gene<-
  factor(up_regulated_sort$gene_name,
         levels = unique(up_regulated_sort$gene_name))
up_regulated_sort$gene <-
  fct_rev(up_regulated_sort$gene)
down_regulated_sort$gene <-
  factor(down_regulated_sort$gene_name,
         levels = unique(down_regulated_sort$gene_name))
down_regulated_sort$gene <-
  fct_rev(down_regulated_sort$gene)

brain_FC_up <- data.frame(up_regulated_sort$gene, up_regulated_sort$tissue, up_regulated_sort$logFC)
brain_FC_down <- data.frame(down_regulated_sort$gene, down_regulated_sort$tissue, down_regulated_sort$logFC)

#remove files
remove(down_regulated_sort,
  down_regulated,
  up_regulated,
  up_regulated_sort, 
  treatment_vs_control
)

# kidney
tissue = "Kidney"
treatment_vs_control <- read.table(
  paste0(
    "../../../results/",
    tool,
    "/DEGs/",
    treatment,
    "_",
    tissue,
    "_gene_DEGs_FDRq1.00.txt",
    sep = ""
  ),
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)

up_regulated <- subset(
  treatment_vs_control,
  treatment_vs_control$adj.P.Val <= 0.05 &
    treatment_vs_control$logFC > 0
)
down_regulated <- subset(
  treatment_vs_control,
  treatment_vs_control$adj.P.Val <= 0.05 &
    treatment_vs_control$logFC <= 0
)

# sort by log2FC and adjusted p-val
up_regulated_sort <-
  up_regulated[order(-up_regulated$logFC, up_regulated$adj.P.Val), ]
down_regulated_sort <-
  down_regulated[order(down_regulated$logFC, down_regulated$adj.P.Val), ]
up_regulated_sort$tissue <- c("Kidney")
down_regulated_sort$tissue <- c("Kidney")


# lock in gene order which is sorted by adjusted p-value
up_regulated_sort$gene<-
  factor(up_regulated_sort$gene_name,
         levels = unique(up_regulated_sort$gene_name))
up_regulated_sort$gene <-
  fct_rev(up_regulated_sort$gene)
down_regulated_sort$gene <-
  factor(down_regulated_sort$gene_name,
         levels = unique(down_regulated_sort$gene_name))
down_regulated_sort$gene <-
  fct_rev(down_regulated_sort$gene)

kidney_FC_up <- data.frame(up_regulated_sort$gene, up_regulated_sort$tissue, up_regulated_sort$logFC)
kidney_FC_down <- data.frame(down_regulated_sort$gene, down_regulated_sort$tissue, down_regulated_sort$logFC)

#remove files
remove(down_regulated_sort,
  down_regulated,
  up_regulated,
  up_regulated_sort,
  treatment_vs_control
)

# rename columns and bind together
names(brain_FC_down)[1:3] <- c('gene', 'tissue', 'logFC')
names(brain_FC_up)[1:3] <- c('gene', 'tissue', 'logFC')
names(kidney_FC_down)[1:3] <- c('gene', 'tissue', 'logFC')
names(kidney_FC_up)[1:3] <- c('gene', 'tissue', 'logFC')

up_FC <- rbind(brain_FC_up, kidney_FC_up)
down_FC <- rbind(brain_FC_down, kidney_FC_down)
```

### Genes in enrichment plot all
## up 
```{r}
up_enrich_df <-
  up_enrich_results_subset[, c("GroupID", "Term", "Description", "Symbols")]
up_enrich_df_melt <- reshape2::melt(up_enrich_df)
up_gene_char <- str_split(up_enrich_df_melt$Symbols, ",")

up_gene_pathway <-
  data.frame(gene = unlist(up_gene_char),
             value = rep(up_enrich_df$Description, lengths(up_gene_char)))
up_pathways_levels <- levels(fct_rev(up_gene_pathway$value))

up_input <- list(
  "Cytokine Signaling in Immune system" = up_gene_char[[1]],
  "regulation of defense response" = up_gene_char[[5]],
  "positive regulation of cell death" = up_gene_char[[8]],
  "TNF weak inducer of apoptosis signaling pathway" = up_gene_char[[11]], 
  "Lipid and atherosclerosis"  = up_gene_char[[15]], 
  "VEGFA-VEGFR2 signaling pathway" = up_gene_char[[16]]
)
up_data <- fromList(up_input)
up_data$gene <- row.names(up_data)
up_data_melt <- reshape2::melt(up_data)


# get the fold change value for those genes
up_df <- merge(up_data_melt, up_FC,
                       by = "gene")
# order by fold change 
up_df <- up_df[order(-up_df$logFC),]
up_df$tissue <- factor(up_df$tissue, levels = c("Kidney", "Brain"))

# subset by GO termes
GO_1 <-
  subset(up_df,
         up_df$variable == "Cytokine Signaling in Immune system"
         & up_df$value == 1)
GO_1 <- subset(GO_1, tissue == "Brain")
GO_1_top <- head(GO_1, 30)

GO_2 <-
  subset(up_df,
         up_df$variable == "regulation of defense response"
         & up_df$value == 1)
GO_2 <- subset(GO_2, tissue == "Brain")
GO_2_top <- head(GO_2, 30)

GO_3 <-
  subset(up_df,
         up_df$variable == "positive regulation of cell death"
         & up_df$value == 1)
GO_3 <- subset(GO_3, tissue == "Brain")
GO_3_top <- head(GO_3, 30)

GO_4 <-
  subset(up_df,
         up_df$variable == "TNF weak inducer of apoptosis signaling pathway"
         & up_df$value == 1)
GO_4 <- subset(GO_4, tissue == "Brain")
GO_4_top <- head(GO_4, 30)

GO_5 <-
  subset(up_df,
         up_df$variable == "Lipid and atherosclerosis"
         & up_df$value == 1)
GO_5 <- subset(GO_5, tissue == "Brain")
GO_5_top <- head(GO_5, 30)

GO_6 <-
  subset(up_df,
         up_df$variable == "VEGFA-VEGFR2 signaling pathway"
         & up_df$value == 1)
GO_6 <- subset(GO_6, tissue == "Brain")
GO_6_top <- head(GO_6, 30)
```
### up 1 heatmap 
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_1_DEG <- subset(up_df, up_df$gene %in% GO_1_top$gene)
GO_1_DEG$gene <- factor(GO_1_DEG$gene,
                          levels = unique(GO_1_top$gene))
GO_1_DEG$gene <- fct_rev(GO_1_DEG$gene)
# heatmap 
up_GO_1_lcpm_heat <- ggplot(data = GO_1_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "white",
    mid = "red",
    high = "maroon",
    midpoint = 4.5,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 3, 6, 9),
    limits = c(0,9),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0.2, 0.5, 0.2, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("Cytokine Signaling in 
Immune system") 
up_GO_1_lcpm_heat
```

### up 2 heatmap 
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_2_DEG <- subset(up_df, up_df$gene %in% GO_2_top$gene)
GO_2_DEG$gene <- factor(GO_2_DEG$gene,
                          levels = unique(GO_2_top$gene))
GO_2_DEG$gene <- fct_rev(GO_2_DEG$gene)
# heatmap 
up_GO_2_lcpm_heat <- ggplot(data = GO_2_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "white",
    mid = "red",
    high = "maroon",
    midpoint = 4.5,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 3, 6, 9),
    limits = c(0,9),
    name = expression(log[2](FC))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0.2, 0.5, 0.2, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("regulation of 
defense response") 
up_GO_2_lcpm_heat
```
### up 3 heatmap 
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_3_DEG <- subset(up_df, up_df$gene %in% GO_3_top$gene)
GO_3_DEG$gene <- factor(GO_3_DEG$gene,
                          levels = unique(GO_3_top$gene))
GO_3_DEG$gene <- fct_rev(GO_3_DEG$gene)
# heatmap 
up_GO_3_lcpm_heat <- ggplot(data = GO_3_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "white",
    mid = "red",
    high = "maroon",
    midpoint = 4.5,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 3, 6, 9),
    limits = c(0,9),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0.2, 0.5, 0.2, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("positive regulation 
of cell death") 
up_GO_3_lcpm_heat
```

### up 4 heatmap 
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_4_DEG <- subset(up_df, up_df$gene %in% GO_4_top$gene)
GO_4_DEG$gene <- factor(GO_4_DEG$gene,
                          levels = unique(GO_4_top$gene))
GO_4_DEG$gene <- fct_rev(GO_4_DEG$gene)
# heatmap 
up_GO_4_lcpm_heat <- ggplot(data = GO_4_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "white",
    mid = "red",
    high = "maroon",
    midpoint = 4.5,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 3, 6, 9),
    limits = c(0,9),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0.2, 0.5, 0.2, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("TNF-related weak 
inducer of apoptosis
signaling pathway") 
up_GO_4_lcpm_heat
```

### up 5 heatmap 
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_5_DEG <- subset(up_df, up_df$gene %in% GO_5_top$gene)
GO_5_DEG$gene <- factor(GO_5_DEG$gene,
                          levels = unique(GO_5_top$gene))
GO_5_DEG$gene <- fct_rev(GO_5_DEG$gene)
# heatmap 
up_GO_5_lcpm_heat <- ggplot(data = GO_5_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "white",
    mid = "red",
    high = "maroon",
    midpoint = 4.5,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 3, 6, 9),
    limits = c(0,9),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0.2, 0.5, 0.2, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("Lipid and 
atherosclerosis") 
up_GO_5_lcpm_heat
```
### up 6 heatmap 
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_6_DEG <- subset(up_df, up_df$gene %in% GO_6_top$gene)
GO_6_DEG$gene <- factor(GO_6_DEG$gene,
                          levels = unique(GO_6_top$gene))
GO_6_DEG$gene <- fct_rev(GO_6_DEG$gene)
# heatmap 
up_GO_6_lcpm_heat <- ggplot(data = GO_6_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "white",
    mid = "red",
    high = "maroon",
    midpoint = 4.5,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 3, 6, 9),
    limits = c(0,9),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0.2, 0.5, 0.2, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("VEGFA-VEGFR2 
signaling pathway") 
up_GO_6_lcpm_heat
```
## up heatmap legend
```{r}
legend_data_up <- ggplot(data = GO_2_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "white",
    mid = "red",
    high = "maroon",
    midpoint = 4.5,
    space = "rgb",
    guide = "colourbar",
    breaks = c(0, 3, 6, 9),
    limits = c(0,9),
    name = expression(log[2](FC))
  ) +
  theme(
    legend.position = "top", 
    legend.text = element_text(size =6), 
    legend.title = element_text (size = 6), 
    legend.margin=margin(0,0,1,0),
  #  legend.direction = "horizontal", 
    legend.key.size = unit(0.5, "cm"))

legend_data_up <- addSmallLegend(legend_data_up)

leg_up <- get_legend(legend_data_up)
# Convert to a ggplot and print
legend_tpm_bar_up <- as_ggplot(leg_up)
legend_tpm_bar_up
```
## down 
* [1] angiogenesis                                             
 [2] hematopoietic or lymphoid organ development              
 [3] Hematopoietic stem cell differentiation                  
* [4] lymph vessel development                                 
* [5] endothelium development                                  
 [6] regulation of cytoskeleton organization                  
 [7] Epithelial to mesenchymal transition in colorectal cancer
* [8] RAC1 GTPase cycle                                        
 [9] regulation of protein binding                            
[10] cellular response to growth factor stimulus              
* [11] phagocytosis 
```{r}
down_enrich_df <-
  down_enrich_results_subset[, c("GroupID", "Term", "Description", "Symbols")]
down_enrich_df_melt <- reshape2::melt(down_enrich_df)
down_gene_char <- str_split(down_enrich_df_melt$Symbols, ",")

down_gene_pathway <-
  data.frame(gene = unlist(down_gene_char),
             value = rep(down_enrich_df$Description, lengths(down_gene_char)))
down_pathways_levels <- levels(fct_rev(down_gene_pathway$value))

down_input <- list(
  "angiogenesis" = down_gene_char[[1]], 
  "lymph vessel development" = down_gene_char[[4]],
  "endothelium development" = down_gene_char[[5]],
  "RAC1 GTPase cycle" = down_gene_char[[8]],
  "phagocytosis" = down_gene_char[[11]]
)
down_data <- fromList(down_input)
down_data$gene <- row.names(down_data)
down_data_melt <- reshape2::melt(down_data)


# get the fold change value for those genes
down_df <- merge(down_data_melt, down_FC,
                       by = "gene")
# order by fold change 
down_df <- down_df[order(-down_df$logFC),]
down_df$tissue <- factor(down_df$tissue, levels = c("Kidney", "Brain"))

# subset by GO termes
GO_1 <-
  subset(down_df,
         down_df$variable == "angiogenesis"
         & down_df$value == 1)
GO_1 <- subset(GO_1, tissue == "Brain")
GO_1_top <- head(GO_1, 35)

GO_2 <-
  subset(down_df,
         down_df$variable == "lymph vessel development"
         & down_df$value == 1)
GO_2 <- subset(GO_2, tissue == "Brain")
GO_2_top <- head(GO_2, 35)

GO_3 <-
  subset(down_df,
         down_df$variable == "endothelium development"
         & down_df$value == 1)
GO_3 <- subset(GO_3, tissue == "Brain")
GO_3_top <- head(GO_3, 35)

GO_4 <-
  subset(down_df,
         down_df$variable == "RAC1 GTPase cycle"
         & down_df$value == 1)
GO_4 <- subset(GO_4, tissue == "Brain")
GO_4_top <- head(GO_4, 35)

GO_5 <-
  subset(down_df,
         down_df$variable == "phagocytosis"
         & down_df$value == 1)
GO_5 <- subset(GO_5, tissue == "Brain")
GO_5_top <- head(GO_5, 35)
```

### down 1 heatmap 
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_1_DEG <- subset(down_df, down_df$gene %in% GO_1_top$gene)
GO_1_DEG$gene <- factor(GO_1_DEG$gene,
                          levels = unique(GO_1_top$gene))
#GO_1_DEG$gene <- fct_rev(GO_1_DEG$gene)
# heatmap 
down_GO_1_lcpm_heat <- ggplot(data = GO_1_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "navyblue",
    mid = "blue",
    high = "white",
    midpoint = -3,
    space = "rgb",
    guide = "colourbar",
    breaks = c(-6, -4, -2,  0),
    limits = c(-6,0),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0.2, 0.5, 0.2, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("angiogenesis") 
down_GO_1_lcpm_heat
```
### down 2 heatmap 
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_2_DEG <- subset(down_df, down_df$gene %in% GO_2_top$gene)
GO_2_DEG$gene <- factor(GO_2_DEG$gene,
                          levels = unique(GO_2_top$gene))
#GO_2_DEG$gene <- fct_rev(GO_2_DEG$gene)
# heatmap 
down_GO_2_lcpm_heat <- ggplot(data = GO_2_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "navyblue",
    mid = "blue",
    high = "white",
    midpoint = -3,
    space = "rgb",
    guide = "colourbar",
    breaks = c(-6, -4, -2,  0),
    limits = c(-6,0),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0.2, 0.5, 0.2, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("lymph vessel 
development") 
down_GO_2_lcpm_heat
```
### down 3 heatmap 
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_3_DEG <- subset(down_df, down_df$gene %in% GO_3_top$gene)
GO_3_DEG$gene <- factor(GO_3_DEG$gene,
                          levels = unique(GO_3_top$gene))
#GO_3_DEG$gene <- fct_rev(GO_3_DEG$gene)
# heatmap 
down_GO_3_lcpm_heat <- ggplot(data = GO_3_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "navyblue",
    mid = "blue",
    high = "white",
    midpoint = -3,
    space = "rgb",
    guide = "colourbar",
    breaks = c(-6, -4, -2,  0),
    limits = c(-6,0),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0.2, 0.5, 0.2, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("endothelium 
development") 
down_GO_3_lcpm_heat
```
### down 4 heatmap 
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_4_DEG <- subset(down_df, down_df$gene %in% GO_4_top$gene)
GO_4_DEG$gene <- factor(GO_4_DEG$gene,
                          levels = unique(GO_4_top$gene))
#GO_4_DEG$gene <- fct_rev(GO_4_DEG$gene)
# heatmap 
down_GO_4_lcpm_heat <- ggplot(data = GO_4_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "navyblue",
    mid = "blue",
    high = "white",
    midpoint = -3,
    space = "rgb",
    guide = "colourbar",
    breaks = c(-6, -4, -2,  0),
    limits = c(-6,0),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0.2, 0.5, 0.2, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("RAC1 GTPase 
cycle") 
down_GO_4_lcpm_heat
```
### down 5 heatmap 
```{r}
# lock in gene order which is sorted by adjusted p-value
GO_5_DEG <- subset(down_df, down_df$gene %in% GO_5_top$gene)
GO_5_DEG$gene <- factor(GO_5_DEG$gene,
                          levels = unique(GO_5_top$gene))
#GO_5_DEG$gene <- fct_rev(GO_5_DEG$gene)
# heatmap 
down_GO_5_lcpm_heat <- ggplot(data = GO_5_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "navyblue",
    mid = "blue",
    high = "white",
    midpoint = -3,
    space = "rgb",
    guide = "colourbar",
    breaks = c(-6, -4, -2,  0),
    limits = c(-6,0),
    name = expression(log[2](CPM))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 6),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = margin(0.2, 0.5, 0.2, 0.2, "cm"), 
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.title.x=element_blank(),
    axis.text.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    ggtitle("phagocytosis") 
down_GO_5_lcpm_heat
```

## down heatmap legend
```{r}
# heatmap 
legend_data_down <- ggplot(data = GO_2_DEG, aes(x = tissue, y = gene)) +
  geom_tile(aes(fill = logFC)) +
  ggplot2::facet_grid(~ tissue, scales = "free", switch = "both") +
  scale_fill_gradient2(
    low = "navyblue",
    mid = "blue",
    high = "white",
    midpoint = -3,
    space = "rgb",
    guide = "colourbar",
    breaks = c(-6,-4,-2,0),
    limits = c(-6,0),
    name = expression(log[2](FC))
  ) +
  scale_x_discrete(expand = c(0,0)) +
  theme(
    legend.position = "top", 
    legend.text = element_text(size =6), 
    legend.title = element_text (size = 6), 
    legend.margin=margin(0,0,1,0),
   # legend.direction = "vertical", 
    legend.key.size = unit(0.5, "cm"))

legend_data_down <- addSmallLegend(legend_data_down)

leg_down <- get_legend(legend_data_down)
# Convert to a ggplot and print
legend_tpm_bar_down <- as_ggplot(leg_down)
legend_tpm_bar_down
```


# Combine plot 
```{r}

row1 <- ggarrange(
  up_and_down_enrich_plot, 
  labels = c("a)"),
  font.label = list(size = 8)
  )


row2 <-
  ggarrange(
    NULL, 
    legend_tpm_bar_up,
    NULL,
    NULL,
    NULL,
    ncol = 5,
    labels = c("up-regulated", ""),
    widths = c(1,0.15, 1, 1,1),
    font.label = list(size = 7, color = "red")
  )

row3 <- ggarrange(
  up_GO_1_lcpm_heat, 
  up_GO_2_lcpm_heat, 
  up_GO_4_lcpm_heat, 
  up_GO_5_lcpm_heat, 
  nrow = 1, 
  ncol = 4,
  labels = c("b)", "c)", "d)", "e)"),
  font.label = list(size = 8)
)

row4 <-
  ggarrange(
    NULL, 
    legend_tpm_bar_down,
    NULL,
    NULL,
    NULL,
    ncol = 5,
    labels = c("down-regulated", ""),
    widths = c(1,0.15, 1, 1,1),
    font.label = list(size = 7, color = "blue")
  )

row5 <- ggarrange(
  down_GO_1_lcpm_heat, 
  down_GO_2_lcpm_heat, 
  down_GO_4_lcpm_heat, 
  down_GO_5_lcpm_heat, 
  nrow = 1, 
  ncol = 4,
  labels = c("f)", "g)", "h)", "i)"),
  font.label = list(size = 8)
)

together <- ggarrange(row1, row2, row3, row4, row5, nrow = 5, heights = c(1.35, 0.125, 1.45, 0.125, .70))
together
path <- paste0("../../../results/manuscript_figures/Supplemental_figure_5_",
               "shared_DEGs_brain_kidney",
               "_heatmaps_v2")
saveToPDF(paste0(path, ".pdf"), width = 7.08, height = 8.66)
```
# session info
```{r}
```
