---
title: "Clinical stats"
author: "Kimberly Olney & Kennedi Todd"
date: "October 1st 2022"
output: 
  pdf_document: default
  html_document:
    df_print: paged
---
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = ".")
```

```{r, message = FALSE, warning = FALSE, echo = FALSE, include = FALSE}
library(ggplot2)
library(ggpubr)
library(purrr)
library(car)
library(reshape)
library(ggforce)
library("knitr")
```

```{r}
saveToPDF <- function(...) {
  d = dev.copy(pdf, ...)
  dev.off(d)
}
```
# Read in clinical data  

```{r read_data, cache=FALSE}
metadata <-
  read.delim((
    "/research/labs/neurology/fryer/projects/sepsis/pig/LPS/metadata.tsv"
  ),
  header = TRUE,
  sep = "\t"
  )
tissue <- c("Brain")
# subset for tissue 
metadata <- metadata[metadata$tissue == tissue, ]

# remove pigs 9 and 13
metadata <- metadata[metadata$pig_id != "9", ]
metadata <- metadata[metadata$pig_id != "13", ]
metadata <- metadata[metadata$blood_group != "BB", ]
metadata <- metadata[!duplicated(metadata[5]),]
```
# Create bar plots for catrogical variables 
```{r bar_plot, cache=FALSE}
# create bar plot showing number of samples for each end point criteria
df <- data.frame(metadata$endpoint)# create dataframe of endpoint criteria 
colnames(df) <- c("endpoint") # rename column 
bar_plot_brain <- ggplot(df, aes(endpoint, ..count..)) + 
  geom_bar(aes(fill = endpoint), position = "dodge") + 
  theme(#axis.title.x=element_blank(),
  axis.text.x=element_blank(),
  axis.ticks.x=element_blank()) + 
  ggtitle("Endpoint summary") +
  scale_y_continuous(breaks=seq(0,6,1)) # Y axis ticks labeling by 1
bar_plot_brain
```
# Test for differences between the two groups (contorls and LPS) for continuous variables
t-test to determine if there is a significant difference between the means of two group
first we determine if the assumptions of a t-test are met. If not, we will log-transform the data and redetermine if the assumptions are met. 
If the assumptions are still not met, we will use a  nonparametric Wilcoxon rank-sum test. 
A Wilcoxon rank-sum test or as known as a Mann Whitney Wilcoxon Test, is used to test whether two groups have the same distibution. 

two sample t-test
    For hypotheses about the difference between the means of two populations. 

First state the null and alternative hypotheses
Ho: Ua - Ub = 0 
    The mean body weight, mean dose is ml per hour, mean mins from injection to sacrifice are the same between controls and LPS pigs. 
 Ho: Ua - Ub != 0 
     The mean body weight, mean dose is ml per hour, mean mins from injection to sacrifice are NOT the same between controls and LPS pigs. 
 the test is carried out independently for each variable being tested. 


testing for normality
calculate variance and standard deviation (sd)
welch's t-test (does not assume equal variance but does assume normality)
wilcoxon/Mann-whit test for differences in distributions. (does not test the mean or sd) 

Levene Test - homogeneity of variance across groups.
homogeneity assumption: the population variances of the dependent variable must equal for all groups. 
Can ignore this assumption if roughly equal sample sizes for each group.
```{r}
variable_names <- c("weight_kg", 
"dose_ml_per_hr", 
"duration_min", 
"start_heart_rate", 
"end_heart_rate", 
"start_temp", 
"end_temp", 
"start_MAP", 
"end_MAP", 
"age_days")

tests <- c("control_mean","LPS_mean","control_stdv","LPS_stdv",
           "control_shapiro_pval","LPS_shapiro_pval","levene","t_test_t",
           "t_test_pval","wilcoxon_pval")
clinical_stats <- matrix(0,
                         nrow = length(variable_names),
                         ncol = length(tests),
                         dimnames = list(variable_names, tests))
clinical_stats <- as.data.frame(clinical_stats)
```

# Run tests
```{r}
# Loop through variables
for (i in 1:length(variable_names)) {

  # Subset values by group
  values <- as.numeric(metadata[, variable_names[i]])
  group <- factor(metadata$group)
  control_values <- metadata[metadata$group == "Control", variable_names[i]]
  LPS_values <- metadata[metadata$group == "LPS", variable_names[i]]
  
  # Welch two sample t-test
  t_test <- t.test(control_values, LPS_values)
  t_test_t <- t_test$statistic
  t_test_pval <- t_test$p.value
  
  # Add to table
  clinical_stats[i, "control_mean"] <- round(as.numeric(t_test$estimate[1]),4)
  clinical_stats[i, "LPS_mean"] <- round(as.numeric(t_test$estimate[2]),4)
  clinical_stats[i, "t_test_t"] <- round(t_test_t,4)
  clinical_stats[i, "t_test_pval"] <- round(t_test_pval,4)
  
  # Shapiro-Wilk test, normality test
  control_shapiro <- shapiro.test(control_values)
  LPS_shapiro <- shapiro.test(LPS_values)
  
  # Add to table
  clinical_stats[i, "control_shapiro_pval"] <- round(control_shapiro$p.value,4)
  clinical_stats[i, "LPS_shapiro_pval"] <- round(LPS_shapiro$p.value,4)
  
  # Variance and standard deviation
  control_var <- var(control_values)
  LPS_var <- var(LPS_values)
 
  # Add to table 
  clinical_stats[i, "control_stdv"] <- round(sqrt(control_var),4)
  clinical_stats[i, "LPS_stdv"] <- round(sqrt(LPS_var),4)
  
  # Two-sample Wilcoxon / Mann-Whitney
  wil <- wilcox.test(control_values, LPS_values)
  
  # Add to table
  clinical_stats[i, "wilcoxon_pval"] <- round(wil$p.value, 4)
  
  # Levene's test
  levTest <- leveneTest(values ~ group)
  lev <- levTest$`Pr(>F)`
  lev_naomit <- as.numeric(na.omit(lev))
  
  # Add to table
  clinical_stats[i, "levene"] <- round(lev_naomit, 4)
}
```

# Output clinical stats table
```{r}
write.table(clinical_stats,
            "../../results/clinical_stats/clinincal_stats.tsv",
            quote = FALSE,
            sep = "\t")
```
# Violin of heart rate and body temperature 
```{r}
#-- heart rate
hr_df <- metadata[c("group", "start_heart_rate", "end_heart_rate")]
hr_df <- melt(hr_df)
supp.labs <- c("Starting heart rate", "Final heart rate")
names(supp.labs) <- c("start_heart_rate", "end_heart_rate")

hr_plot <- ggplot(hr_df, aes(x = group, y = value, color = group))  +
    geom_violin() + scale_color_manual(values = c("black", "purple")) +
    geom_boxplot(width = 0.1, outlier.size = -1, outlier.shape = NA) +
    geom_sina(aes(shape = factor(group)),
                size = 1.5) +
    ggplot2::facet_grid(~ variable, labeller = labeller(variable = supp.labs)) +
    theme_bw() +
  theme(
    plot.margin = margin(0.1, 0.2, 0, 0.2, "cm"), 
    strip.text = element_text(size = 6.5, margin = margin(1,0,1.2,0)),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.text.x=element_text(size = 6),
    axis.title.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_text(size = 6), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    labs(
    y = "beats per minute"
  ) + ggtitle("Heart rate") + scale_shape_manual(
      values = c(15, 19)) +
  scale_y_continuous(breaks = seq(60, 230, by =20), limits = c(60, 250)) 
hr_plot
hr_plot <- hr_plot +  theme(panel.spacing.x = unit(0.7, "lines"))
```

```{r}
#-- Temp
temp_df <- metadata[c("group", "start_temp", "end_temp")]
temp_df <- melt(temp_df)
supp.labs <- c("Starting temperature", "Final temperature")
names(supp.labs) <- c("start_temp", "end_temp")

temp_plot <- ggplot(temp_df, aes(x = group, y = value, color = group))  +
    geom_violin() + scale_color_manual(values = c("black", "purple")) +
    geom_boxplot(width = 0.1, outlier.size = -1, outlier.shape = NA) +
    geom_sina(aes(shape = factor(group)),
                size = 1.5) +
    ggplot2::facet_grid(~ variable, labeller = labeller(variable = supp.labs)) +
    theme_bw() +
  theme(
    plot.margin = margin(0.1, 0.2, 0, 0.2, "cm"), 
    strip.text = element_text(size = 6.5, margin = margin(1,0,1.2,0)),
    legend.position = "none",
    panel.border = element_blank(),
    panel.background = element_blank(),
    panel.spacing = unit(0,'lines'), 
    plot.title = element_text(size = 7, vjust = -1, hjust = 0),
    axis.text.x=element_text(size = 6),
    axis.title.x = element_blank(), 
    axis.ticks.x=element_blank(), 
    axis.title.y = element_text(size = 6), 
    axis.text.y = element_text(size = 6, margin = margin(r = -2)),
    axis.ticks.y=element_blank()) +
    labs(
    y = "Fahrenheit"
  ) + ggtitle("Temperature") + scale_shape_manual(
      values = c(15, 19)) +
  scale_y_continuous(breaks = seq(95, 115, by =5), limits = c(95, 118)) 
 #   sec.axis = sec_axis(~ . *9/5 +32)
temp_plot
temp_plot <- temp_plot +  theme(panel.spacing.x = unit(0.7, "lines"))
```

# Combine plot 
```{r}

row1 <- 
    ggarrange(
    hr_plot,
    temp_plot,
    ncol = 2,
    nrow = 1, 
    labels = c("a)", "b)"),
    font.label = list(size = 8)
  )

together <- ggarrange(row1, NULL, nrow = 2, heights = c(1, 3))
together
path <- paste0("../../results/manuscript_figures/Figure_1_",
               "clinical_and_pathology")
saveToPDF(paste0(path, ".pdf"), width = 7.08, height = 8.66)
```
# Plot everything 
```{r violin_jitter, cache=FALSE}
pheno_df <- data.frame(metadata$weight_kg, metadata$dose_ml_per_hr, metadata$duration_min, metadata$start_heart_rate, metadata$end_heart_rate, metadata$start_temp, metadata$end_temp, metadata$start_MAP, metadata$end_MAP, metadata$age_days)
c_pheno_df <- colnames(df)
labels <-
  c(
    "Weight (kg)",
    "Dose mL per hour",
    "Mins from injection till sacrifice", 
    "Initial heart rate", 
    "Final heart rate", 
    "Initial body temperature",
    "Final body temperature", 
    "Initial MAP", 
    "Final  MAP", 
    "Days old" 
  )
vio_Func_ttest <- function(j, i, k) {
  ggplot(metadata, aes(x = group, y = i, color = group)) + ylab(j)  +
    geom_violin() + scale_color_manual(values = c("black", "purple")) +
    geom_boxplot(width = 0.1, outlier.size = -1, outlier.shape = NA) + 
    stat_compare_means(method = "t.test", size = 4, label.x = 1.25) +
    geom_sina(aes(shape = factor(group)),
                size = 3) +
    theme(legend.position = "none") + ggtitle(k) + ylab(k) + scale_shape_manual(
      values = c(14, 10)) +
    theme(
      axis.title.x = element_text(size = 12),
      axis.text.x =
        element_text(size = 12)
    ) +
    theme(
      axis.title.y = element_text(size = 12),
      axis.text.y =
        element_text(size = 12) 
    ) +
    theme(
      title = element_text(size = 12)
    )
}
all_plots <- Map(vio_Func_ttest, i = pheno_df, j = c_pheno_df, k = labels)
all_plots
plotnames = imap(all_plots, ~paste0("", .y, "_ttest.pdf")) %>%
  flatten()
plotnames
walk2(plotnames, all_plots, ~ggsave(filename = .x, plot = .y, 
                                    height = 4, width = 4))
# Wilcox
vio_Func_wilcox <- function(j, i, k) {
  ggplot(metadata, aes(x = group, y = i, color = group)) + ylab(j)  +
    geom_violin() + scale_color_manual(values = c("black", "purple")) +
    geom_boxplot(width = 0.1, outlier.size = -1, outlier.shape = NA) + 
    stat_compare_means(method = "wilcox.test", paired = FALSE, size = 4, label.x = 1.25) +
    geom_sina(aes(shape = factor(group)),
                size = 3) +
    theme(legend.position = "none") + ggtitle(k) + ylab(k) + scale_shape_manual(
      values = c(14, 10)) +
    theme(
      axis.title.x = element_text(size = 12),
      axis.text.x =
        element_text(size = 12)
    ) +
    theme(
      axis.title.y = element_text(size = 12),
      axis.text.y =
        element_text(size = 12) 
    ) +
    theme(
      title = element_text(size = 12)
    )
}
# Map: iterates through items in Meta (which is a list of dataframes) 
# and iterates through the names of the items in Meta simultaneously
all_plots <- Map(vio_Func_wilcox, i = pheno_df, j = c_pheno_df, k = labels)
all_plots
plotnames = imap(all_plots, 
                 ~paste0("../../results/clinical_stats/", 
                         .y, "_wilcox.pdf")) %>%
  flatten()
plotnames
walk2(plotnames, all_plots, ~ggsave(filename = .x, plot = .y, 
                                    height = 4, width = 4))
```