---
title: "Get human ortholog for pig gene names"
author: "Kimberly Olney"
date: "7/26/2022"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: inline
---


```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = ".")
```

```{r, message = FALSE, warning = FALSE, echo = FALSE, include = FALSE}
library(geneName)
library(plotly)
library(tidyverse)
library(htmlwidgets)
library(ggvenn)
library(dplyr)
library(gprofiler2)
library(stats)
```

# Read in DEG tables
```{r}
pig_brain_FDRq1.00 <-
  read.delim("../../results/star/DEGs/LPS_Brain_gene_DEGs_FDRq1.00.txt")

pig_kidney_FDRq1.00 <-
  read.delim("../../results/star/DEGs/LPS_Kidney_gene_DEGs_FDRq1.00.txt")

pig_blood_FDRq1.00 <-
  read.delim("../../results/star/DEGs/LPS_Blood_gene_DEGs_FDRq1.00.txt")
```

# Orthologs
Convert mouse and pig gene_ids to human gene symbols 
https://cran.r-project.org/web/packages/gprofiler2/vignettes/gprofiler2.html
```{r}
human_pig_brain_orth <-
  gorth(
    query = pig_brain_FDRq1.00$gene_id,
    source_organism = "sscrofa",
    target_organism = "hsapiens",
    mthreshold = Inf,
    filter_na = TRUE,
    numeric_ns = "ENTREZGENE_ACC"
  )

human_pig_kidney_orth <-
  gorth(
    query = pig_brain_FDRq1.00$gene_id,
    source_organism = "sscrofa",
    target_organism = "hsapiens",
    mthreshold = Inf,
    filter_na = TRUE,
    numeric_ns = "ENTREZGENE_ACC"
  )

human_pig_blood_orth <-
  gorth(
    query = pig_brain_FDRq1.00$gene_id,
    source_organism = "sscrofa",
    target_organism = "hsapiens",
    mthreshold = Inf,
    filter_na = TRUE,
    numeric_ns = "ENTREZGENE_ACC"
  )
```
# Merge ortholog name with the species gene id
```{r}
# add column for downstream merging
human_pig_orth$gene_id <- human_pig_orth$input
pig_brain_FDRq1.00_orth <-
  merge(pig_brain_FDRq1.00, human_pig_orth, by = "gene_id")
pig_brain_FDRq0.05_orth <-
  merge(pig_brain_FDRq0.05, human_pig_orth, by = "gene_id")
# remove duplicates and save as a separate table for down stream investigation. 
pig_n_occur <- data.frame(table(pig_brain_FDRq1.00_orth$ortholog_name))
names(pig_n_occur)[names(pig_n_occur) == 'Var1'] <- 'ortholog_name'
names(pig_n_occur)[names(pig_n_occur) == 'Freq'] <- 'ortholog_name_frequency'
pig_orth_gene_freq <- merge(pig_brain_FDRq1.00_orth, pig_n_occur, by = "ortholog_name", all.x = TRUE)

