---
title: "LPS mouse and pig DEG comparison and correlation"
author: "Kimberly Olney"
date: "7/26/2022"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = ".")
```

```{r, message = FALSE, warning = FALSE, echo = FALSE, include = FALSE}
library(geneName)
library(plotly)
library(tidyverse)
library(htmlwidgets)
library(ggvenn)
library(dplyr)
library(gprofiler2)
library(stats)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE, include = FALSE}
saveToPDF <- function(...) {
  d = dev.copy(pdf, ...)
  dev.off(d)
}
```

# Read in DEG tables
```{r}
Kang <-
  read.delim("../../mouse_comparison/mouse_WT-LPS_over_WT.txt")
# fdrq 0.05
mouse_limma_FDRq0.05 <-
  read.delim("../../mouse_comparison/results/LPS_WT_gene_DEGs_FDRq0.05.txt")
pig_brain_FDRq0.05 <-
  read.delim("../../results/star/DEGs/LPS_Brain_gene_DEGs_FDRq0.05.txt")
# fdrq 1.0
mouse_limma_FDRq1.00 <-
  read.delim("../../mouse_comparison/results/LPS_WT_gene_DEGs_FDRq1.00.txt")
pig_brain_FDRq1.00 <-
  read.delim("../../results/star/DEGs/LPS_Brain_gene_DEGs_FDRq1.00.txt")
```

# Orthologs
Convert mouse and pig gene_ids to human gene symbols 
https://cran.r-project.org/web/packages/gprofiler2/vignettes/gprofiler2.html
```{r}
human_mouse_orth <-
  gorth(
    query = mouse_limma_FDRq1.00$GeneID,
    source_organism = "mmusculus",
    target_organism = "hsapiens",
    mthreshold = Inf,
    filter_na = TRUE, 
    numeric_ns = "ENTREZGENE_ACC"
  )

human_pig_orth <-
  gorth(
    query = pig_brain_FDRq1.00$gene_id,
    source_organism = "sscrofa",
    target_organism = "hsapiens",
    mthreshold = Inf,
    filter_na = TRUE,
    numeric_ns = "ENTREZGENE_ACC"
  )
```
# Merge ortholog name with the species gene id
```{r}
# add column for downstream merging
human_pig_orth$gene_id <- human_pig_orth$input
pig_brain_FDRq1.00_orth <-
  merge(pig_brain_FDRq1.00, human_pig_orth, by = "gene_id")
pig_brain_FDRq0.05_orth <-
  merge(pig_brain_FDRq0.05, human_pig_orth, by = "gene_id")
# remove duplicates and save as a separate table for down stream investigation. 
pig_n_occur <- data.frame(table(pig_brain_FDRq1.00_orth$ortholog_name))
names(pig_n_occur)[names(pig_n_occur) == 'Var1'] <- 'ortholog_name'
names(pig_n_occur)[names(pig_n_occur) == 'Freq'] <- 'ortholog_name_frequency'
pig_orth_gene_freq <- merge(pig_brain_FDRq1.00_orth, pig_n_occur, by = "ortholog_name", all.x = TRUE)

pig_brain_FDRq1.00_orth_1n <- pig_brain_FDRq1.00_orth[pig_brain_FDRq1.00_orth$ortholog_name %in% pig_n_occur$ortholog_name[pig_n_occur$ortholog_name_frequency == 1],]

pig_brain_FDRq1.00_orth_dups <- pig_brain_FDRq1.00_orth[pig_brain_FDRq1.00_orth$ortholog_name %in% pig_n_occur$ortholog_name[pig_n_occur$ortholog_name_frequency > 1],]
#remove(pig_n_occur)

#pig_n_occur <- data.frame(table(pig_brain_FDRq0.05_orth$ortholog_name))
#pig_n_occur[pig_n_occur$ortholog_name_frequency > 1,]
pig_brain_FDRq0.05_orth_1n <- pig_brain_FDRq0.05_orth[pig_brain_FDRq0.05_orth$ortholog_name %in% pig_n_occur$ortholog_name[pig_n_occur$ortholog_name_frequency == 1],]

pig_brain_FDRq0.05_orth_dups <- pig_brain_FDRq0.05_orth[pig_brain_FDRq0.05_orth$ortholog_name %in% pig_n_occur$ortholog_name[pig_n_occur$ortholog_name_frequency > 1],]
# output as data frames
write.table(
  pig_orth_gene_freq,
  "../../mouse_comparison/results/pig_brain_FDRq1.00_orth.txt",
  sep = "\t",
  quote = F,
  row.names = F
)

write.table(
  pig_brain_FDRq1.00_orth_dups,
  "../../mouse_comparison/results/pig_brain_FDRq1.00_orth_dups.txt",
  sep = "\t",
  quote = F,
  row.names = F
)

write.table(
  pig_brain_FDRq1.00_orth_1n,
  "../../mouse_comparison/results/pig_brain_FDRq1.00_orth_1n.txt",
  sep = "\t",
  quote = F,
  row.names = F
)
```

```{r}
human_mouse_orth$GeneID <- human_mouse_orth$input
mouse_limma_FDRq1.00_orth <-
  merge(mouse_limma_FDRq1.00, human_mouse_orth, by = "GeneID")
mouse_limma_FDRq0.05_orth <-
  merge(mouse_limma_FDRq0.05, human_mouse_orth, by = "GeneID")

mouse_n_occur <- data.frame(table(mouse_limma_FDRq1.00_orth$ortholog_name))
names(mouse_n_occur)[names(mouse_n_occur) == 'Var1'] <- 'ortholog_name'
names(mouse_n_occur)[names(mouse_n_occur) == 'Freq'] <- 'ortholog_name_frequency'
mouse_orth_gene_freq <- merge(mouse_limma_FDRq1.00_orth, mouse_n_occur, by = "ortholog_name", all.x = TRUE)

mouse_limma_FDRq1.00_orth_1n <- mouse_limma_FDRq1.00_orth[mouse_limma_FDRq1.00_orth$ortholog_name %in% mouse_n_occur$ortholog_name[mouse_n_occur$ortholog_name_frequency == 1],]

mouse_limma_FDRq1.00_orth_dups <- mouse_limma_FDRq1.00_orth[mouse_limma_FDRq1.00_orth$ortholog_name %in% mouse_n_occur$ortholog_name[mouse_n_occur$ortholog_name_frequency > 1],]


mouse_limma_FDRq0.05_orth_1n <- mouse_limma_FDRq0.05_orth[mouse_limma_FDRq0.05_orth$ortholog_name %in% mouse_n_occur$ortholog_name[mouse_n_occur$ortholog_name_frequency == 1],]

mouse_limma_FDRq0.05_orth_dups <- mouse_limma_FDRq0.05_orth[mouse_limma_FDRq0.05_orth$ortholog_name %in% mouse_n_occur$ortholog_name[mouse_n_occur$ortholog_name_frequency > 1],]

# output as data frames
write.table(
  mouse_orth_gene_freq,
  "../../mouse_comparison/results/mouse_limma_FDRq1.00_orth.txt",
  sep = "\t",
  quote = F,
  row.names = F
)

write.table(
  mouse_limma_FDRq1.00_orth_dups,
  "../../mouse_comparison/results/mouse_limma_FDRq1.00_orth_dups.txt",
  sep = "\t",
  quote = F,
  row.names = F
)

write.table(
  mouse_limma_FDRq1.00_orth_1n,
  "../../mouse_comparison/results/mouse_limma_FDRq1.00_orth_1n.txt",
  sep = "\t",
  quote = F,
  row.names = F
)
```

# Find shared and unique DEGs between mouse and pig
```{r}
# rename GeneName to gene_name to merge data sets and find intersections
mouse_and_pg_df <-
  merge(mouse_limma_FDRq1.00_orth_1n, pig_brain_FDRq1.00_orth_1n, by = "ortholog_name")

shared_gene_name <-
  intersect(mouse_limma_FDRq0.05_orth_1n$ortholog_name,
            pig_brain_FDRq0.05_orth_1n$ortholog_name)
uniqueToMouse_gene_name <-
  setdiff(mouse_limma_FDRq0.05_orth_1n$ortholog_name,
          pig_brain_FDRq0.05_orth_1n$ortholog_name)
uniqueToPig_gene_name <-
  setdiff(pig_brain_FDRq0.05_orth_1n$ortholog_name,
          mouse_limma_FDRq0.05_orth_1n$ortholog_name)

shared <-
  mouse_and_pg_df[mouse_and_pg_df$ortholog_name %in% shared_gene_name,]
uniqueToMouse <-
  mouse_and_pg_df[mouse_and_pg_df$ortholog_name %in% uniqueToMouse_gene_name,]
uniqueToPig <-
  mouse_and_pg_df[mouse_and_pg_df$ortholog_name %in% uniqueToPig_gene_name,]


# Venn diagram to see total DEGs between mouse and pig
x = list(mouse = mouse_limma_FDRq0.05_orth$ortholog_name,
         pig = pig_brain_FDRq0.05_orth$ortholog_name)

mouse_pig_venn <- ggvenn(
  x,
  fill_color = c("#EFC000FF", "purple"),
  stroke_size = 2,
  set_name_size = 6
)
mouse_pig_venn
```

```{r,  message = FALSE, warning = FALSE, echo = FALSE, include = FALSE}
mouse_pig_venn
path <-
  paste0("../../mouse_comparison/results/Venn_mouse_and_pig_brain")
saveToPDF(paste0(path, ".pdf"), width = 8, height = 8)
```

# Output shared and unique DEGs as tables
```{r}
#output table
write.table(
  shared_gene_name,
  "../../mouse_comparison/results/mouse_pig_shared_gene_name.txt",
  quote = F,
  row.names = F
)

write.table(
  uniqueToMouse_gene_name,
  "../../mouse_comparison/results/uniqueToMouse_gene_name.txt",
  quote = F,
  row.names = F
)

write.table(
  uniqueToPig_gene_name,
  "../../mouse_comparison/results/uniqueToPig_gene_name.txt",
  quote = F,
  row.names = F
)
```

# Correlation of the log2FC between species 
### Shared DEGs
```{r}
# Shapiro-Wilk normality test
shapiro.test(shared$logFC.x)
shapiro.test(shared$logFC.y)

# correlation test
res <- cor.test(shared$logFC.x, shared$logFC.y,
                method = "spearman")
p_value <- round(res$p.value, 3)
rho_value <- round(res$estimate, 3)

# shared between mouse and pig
p <-
  ggplot(data = shared, aes(
    x = logFC.x,
    y = logFC.y,
    text = paste(ortholog_name)
  )) +
  annotate(
    "rect",
    xmin = 0,
    xmax = max(shared$logFC.x),
    ymin = max(shared$logFC.y),
    ymax = 0,
    fill = "lightpink3",
    alpha = .5
  ) +  annotate(
    "rect",
    xmin = 0,
    xmax = min(shared$logFC.x),
    ymin = 0,
    ymax = min(shared$logFC.y),
    fill = "cadetblue3",
    alpha = .5
  ) +
  geom_abline(color = "gray40") +
  geom_point(size = 2) +
  theme_bw() +
  theme(plot.title = element_text(size = 10)) +
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size = 10),
        axis.text.x = element_text(size = 10)) +
  theme(axis.title.y = element_text(size = 10),
        axis.text.y = element_text(size = 10)) +
  annotate(
    "text",
    x = -2,
    y = 5,
    label = paste0("rho = ", rho_value, "\n",
                   "p = ", p_value),
    parse = TRUE,
    color = "gray29",
    size = 5
  ) + labs(title = "Shared DEGs between mouse and pig",
           x = "mouse",
           y = "pig")

ggplotly(p)
```

### Unique to mouse DEGs
```{r}
shapiro.test(uniqueToMouse$logFC.x)
shapiro.test(uniqueToMouse$logFC.y)

res <- cor.test(uniqueToMouse$logFC.x, uniqueToMouse$logFC.y,
                method = "spearman")
p_value <- round(res$p.value, 3)
rho_value <- round(res$estimate, 3)

p <-
  ggplot(data = uniqueToMouse, aes(
    x = logFC.x,
    y = logFC.y,
    text = paste(gene_name)
  )) +
  annotate(
    "rect",
    xmin = 0,
    xmax = max(uniqueToMouse$logFC.x),
    ymin = max(uniqueToMouse$logFC.y),
    ymax = 0,
    fill = "lightpink3",
    alpha = .5
  ) +  annotate(
    "rect",
    xmin = 0,
    xmax = min(uniqueToMouse$logFC.x),
    ymin = 0,
    ymax = min(uniqueToMouse$logFC.y),
    fill = "cadetblue3",
    alpha = .5
  ) +
  geom_abline(color = "gray40") +
  geom_point(size = 2) +
  theme_bw() +
  theme(plot.title = element_text(size = 10)) +
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size = 10),
        axis.text.x = element_text(size = 10)) +
  theme(axis.title.y = element_text(size = 10),
        axis.text.y = element_text(size = 10)) +
  annotate(
    "text",
    x = -2,
    y = 1.25,
    label = paste0("rho = ", rho_value, "\n",
                   "p = ", p_value),
    parse = TRUE,
    color = "gray29",
    size = 5
  ) + labs(title = "DEGs unique to mouse",
           x = "mouse",
           y = "pig")

ggplotly(p)
```

### Unique to pig DEGs
```{r}
shapiro.test(uniqueToPig$logFC.x)
shapiro.test(uniqueToPig$logFC.y)

res <- cor.test(uniqueToPig$logFC.x, uniqueToPig$logFC.y,
                method = "spearman")
p_value <- round(res$p.value, 3)
rho_value <- round(res$estimate, 3)

p <-
  ggplot(data = uniqueToPig, aes(
    x = logFC.x,
    y = logFC.y,
    text = paste(gene_name)
  )) +
  annotate(
    "rect",
    xmin = 0,
    xmax = max(uniqueToPig$logFC.x),
    ymin = max(uniqueToPig$logFC.y),
    ymax = 0,
    fill = "lightpink3",
    alpha = .5
  ) +  annotate(
    "rect",
    xmin = 0,
    xmax = min(uniqueToPig$logFC.x),
    ymin = 0,
    ymax = min(uniqueToPig$logFC.y),
    fill = "cadetblue3",
    alpha = .5
  ) +
  geom_abline(color = "gray40") +
  geom_point(size = 2) +
  theme_bw() +
  theme(plot.title = element_text(size = 10)) +
  theme(legend.position = "none") +
  theme(axis.title.x = element_text(size = 10),
        axis.text.x = element_text(size = 10)) +
  theme(axis.title.y = element_text(size = 10),
        axis.text.y = element_text(size = 10)) +
  annotate(
    "text",
    x = -.17,
    y = 1.5,
    label = paste0("rho = ", rho_value, "\n",
                   "p = ", p_value),
    parse = TRUE,
    color = "gray29",
    size = 5
  ) +
  labs(title = "DEGs unique to pig",
       x = "mouse",
       y = "pig")
ggplotly(p)
```
# Write output tables
### shared 
```{r}
# difference between fold change
shared$logFC_difference <- (shared$logFC.x - shared$logFC.y)
# add direction information. 
# Is the gene down-regulated or up-regulated in both the mouse and pig data sets
shared$direction_value <- shared$logFC.x * shared$logFC.y

df_shared <- shared %>%
  mutate(direction_status = if_else(shared$direction_value < 0, "opposite", "same"))

# now reformat the whole dataframe to be pretty and legible
df_shared <-
  df_shared[, c(1,2,9,13,21,28,32,40,42)]
# rename columns
data.table::setnames(
  df_shared,
  old = c('GeneID','logFC.x', 'adj.P.Val.x', 'gene_id', 'logFC.y', 'adj.P.Val.y'),
  new = c(
    'mouse_gene_id',
    'mouse_logFC',
    'mouse_adj.P.Val',
    'pig_gene_id',
    'pig_logFC',
    'pig_adj.P.Val'
  )
)

write.table(
  df_shared,
  "../../mouse_comparison/results/df_shared_mouse_pig.txt",
  sep = "\t",
  quote = F,
  row.names = F
)
```
### unique to mouse
```{r}
# difference between fold change
uniqueToMouse$logFC_difference <- (uniqueToMouse$logFC.x - uniqueToMouse$logFC.y)
# add direction information. 
# Is the gene down-regulated or up-regulated in both the mouse and pig data sets
uniqueToMouse$direction_value <- uniqueToMouse$logFC.x * uniqueToMouse$logFC.y

df_uniqueToMouse <- uniqueToMouse %>%
  mutate(direction_status = if_else(uniqueToMouse$direction_value < 0, "opposite", "same"))

# now reformat the whole dataframe to be pretty and legible
df_uniqueToMouse <-
  df_uniqueToMouse[, c(1,2,9,13,21,28,32,40,42)]
# rename columns
data.table::setnames(
  df_uniqueToMouse,
  old = c('GeneID','logFC.x', 'adj.P.Val.x', 'gene_id', 'logFC.y', 'adj.P.Val.y'),
  new = c(
    'mouse_gene_id',
    'mouse_logFC',
    'mouse_adj.P.Val',
    'pig_gene_id',
    'pig_logFC',
    'pig_adj.P.Val'
  )
)

write.table(
  df_uniqueToMouse,
  "../../mouse_comparison/results/df_uniqueToMouse_mouse_pig.txt",
  sep = "\t",
  quote = F,
  row.names = F
)
```
### unique to pig
```{r}
# difference between fold change
uniqueToPig$logFC_difference <- (uniqueToPig$logFC.x - uniqueToPig$logFC.y)
# add direction information. 
# Is the gene down-regulated or up-regulated in both the mouse and pig data sets
uniqueToPig$direction_value <- uniqueToPig$logFC.x * uniqueToPig$logFC.y

df_uniqueToPig <- uniqueToPig %>%
  mutate(direction_status = if_else(uniqueToPig$direction_value < 0, "opposite", "same"))

# now reformat the whole dataframe to be pretty and legible
df_uniqueToPig <-
  df_uniqueToPig[, c(1,2,9,13,21,28,32,40,42)]
# rename columns
data.table::setnames(
  df_uniqueToPig,
  old = c('GeneID','logFC.x', 'adj.P.Val.x', 'gene_id', 'logFC.y', 'adj.P.Val.y'),
  new = c(
    'mouse_gene_id',
    'mouse_logFC',
    'mouse_adj.P.Val',
    'pig_gene_id',
    'pig_logFC',
    'pig_adj.P.Val'
  )
)

write.table(
  df_uniqueToPig,
  "../../mouse_comparison/results/df_uniqueToPig_mouse_pig.txt",
  sep = "\t",
  quote = F,
  row.names = F
)
```
# Save rObjects
```{r}
```
